---
title: "AHgen - Vignette 1, Getting started"
author: "Dr Annie Visser-Quinn, a.visser-quinn@hw.ac.uk"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{AHgen - Vignette 1, Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(
  root.dir = normalizePath('../AHgen')
  )

rm(list = ls()); cat("\014")

library(tidyverse)
library(igraph)

```

```{r, eval = FALSE, include = FALSE}
getwd()
```
<span style="color: red;">Couple of lines for introduction. Describe what Abstraction Hierarchy / Urban System Abstraction Hierarchy (USAH) is and point user in right direction.</span>

<p>

## 1. Prerequisites
- OSMtidy V0.0.21
- R version 3.6.3 or newer; type `version` in the console to check
- R packages `tidyverse`, `igraph`, `sf`, `readxl` and `openxlsx`

To install these packages: `install.packages(c("tidyverse", "igraph", "sf", "readxl", "openxlsx"))`. Packages can be loaded using `library` as required.

## 2. Set-up and functions

To use AHgen, you should save your R script within the *AHgen* directory. Set up your R script by (copy and paste from the code chunk below): (1) clearing the environment and console, (2) setting the working directory to your script directory, and (3) loading the required packages.

```{r, eval = FALSE}
rm(list = ls()); cat("\014") 

setwd(dirname(rstudioapi::getSourceEditorContext()$path))

library(tidyverse)
library(igraph)
library(readxl)
```

<p>

The functions that make up AHgen are located in the *functions* folder. You can load all of them in by running the script in *functions.R* using `source()`. In the code chunk below, `ls()` prints the names of the functions loaded into the environment.
```{r}
source("functions/functions.R")

ls()
```


## 3. Data - Read adjacency matrix
The USAH generic template is represented as an adjacency matrix. <span style="color: red;">Some descriptor on what adjacency matrix is. Points to somewhere else for details. Can modify the matrix or add a different one for different applications.</span>

All inputs for AHgen are located in the folder *inputs*, the adjacency matrix files have the prefix *USAH/adjMat*. Two versions in different formats are available: *.csv*, a plain-text version and *.xlsx*, a user-friendly navigable Excel spreadsheet.

The adjacency matrix can be read in using the functions `read_adjMat_csv()` or `read_adjMat_xlsx()`. The former has only one input argument - the filename. The latter requires the filename as well as the sheet number; this is set to 1 by default.

```{r, eval = FALSE}
# All 3 function calls below import the same generic template as an adjacency matrix
read_adjMat_csv("inputs/USAH_adjMat_2020-05-20.csv")
read_adjMat_xlsx("inputs/USAH_adjMat_2020-05-20.xlsx")
read_adjMat_xlsx("inputs/USAH_adjMat_2020-05-20.xlsx", sheet = 1)
```

<p>

*Note* Previews and summaries of the adjacency matrix are long in R. For this vignette we examine the adjacency matrix using `str(list.len = 10, width = 100, strict.width = "cut")`.

```{r}
da <- read_adjMat_xlsx("inputs/USAH_adjMat_2020-05-20.xlsx")
da %>% str(list.len = 10, width = 100, strict.width = "cut")
```

## 4. Data - Node information
The second input required is the node information - this can also be found in the *inputs* folder. The functions follow the same structure as the *read_adjMat* functions: `read_nodeInfo_csv()` and `read_nodeInfo_xlsx()`.

*Note* Again, this vignette examines the node information data using `str()`

```{r}
dn <- read_nodeInfo_xlsx("inputs/USAH_nodeInfo_2020-05-20.xlsx")
dn %>% str(width = 100, strict.width = "cut")
```

## 5. Conversion to edge list and igraph
Weighting and analysis in AHgen requires an edge list or igraph. <span style="color: red;">Descriptions for edge list and igraph with forward links.</span>


### 5.1 Adjacency matrix to edge list
To convert the adjacency matrix into an edge list use the function `adjMat_to_edgelist()`. There are two input arguments: the adjacency matrix and whether the output should be a list of the levels of the abstraction hierarchy (TRUE or FALSE). 

```{r}
de <- da %>% adjMat_to_edgelist
de

de_list <- da %>% adjMat_to_edgelist(outputList = TRUE)
de_list
```

<p>

The edge list is in a tibble format, it can be navigated and edited like any data.frame or tibble in R.

```{r}
de$layers %>% unique # Unique layer names

de %>% filter(layers == "l1FP_l2VPM") # Filter layers

de %>% filter_all(any_vars(str_detect(str_to_lower(.), "collective")))

de %>% filter_all(any_vars(str_detect(str_to_lower(.), "subway")))

de %>% filter(weight != 1) # None, all are equal weight in generic template
```

<p>

<span style="color: green;">Would the filter piece of code be useful as a function for AHgen?</span>

Use the tidyverse function `bind_rows()` to collapse the list of edge lists.

```{r}
de_list %>% bind_rows()
```

### 5.2 Adjacency matrix to igraph
The adjacency matrix can be converted directly to an igraph using the function `adjMat_to_igraph()`. The function has one input argument, the adjacency matrix.

```{r} 
di <- da %>% adjMat_to_igraph
di
```

The top two lines in above output tell us about the igraph. On the first line, it states that this is an igraph and provides a code name for the graph. The four bits which follow are the code that denotes the type of igraph; the abstraction hierarchy is *U*ndirected, the *N*odes have names and the *W*eight attribute has been set. The fourth bit is whether the graph is bipartite (it is not, hence *-*). 

The second line provides the attributes of the nodes (denoted by v) and edges (denoted by e). The abstraction hierarchy has named nodes *name (v/c)* and weighted edges *weight (e/n)*. The letter after the forward slash tells you the class of this attribute, character and numeric in this case. 

The rest of the output is the names of the edges. See the igraph package for details on how to explore the igraph further. Some simple examples are provided below.

```{r}
V(di)

E(di)

E(di)$weight %>% head(n = 25)

E(di)$layers %>% head(n = 25)
```


### 5.3 Edge list to igraph

The edge list can be converted to an igraph using the function `edgelist_to_igraph()`. The function has two input arguments, an edge list and whether the output should be in list form. 

```{r}
de %>% edgelist_to_igraph(outputList = FALSE)

de %>% edgelist_to_igraph(outputList = TRUE)
```


### 5.4 igraph to edge list

<span style="color: green;">Should AHgen provide a function for this? It's very simple, but for the sake of completeness maybe?</span>

```{r, eval = FALSE}
di %>% get.data.frame %>% as_tibble() %>% select(layers, everything())
```

The walkthrough continues with **Part 2 - Link weighting**.