---
title: "Analyse - Additional functions"
author: 
 - Annie Visser-Quinn & Melissa Bedinger
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: paper
    highlight: haddock
vignette: >
  %\VignetteIndexEntry{5 - Analyse - Additional functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)

rm(list = ls()); cat("\014")

library(knitr)
library(tidyverse)
library(AHgen)
```

<style>
#TOC {
  background: url("images/hex.png");
  background-size: 230px;
  background-position-x: center;
  background-position-y: top;
  padding-top: 280px !important;
  background-repeat: no-repeat;
}
</style>
---

In this vignette, you will find out about analysing Abstraction Hierarchies with additional `AHgen` functions, building on the core functions from [Vignette 4](Vignette-4---Analyse---Core-functions.html):

- Section 1 - Geospatial analysis
- Section 2 - Scenario analysis
- Section 3 - Sensitivity analysis
- Section 4 - Comparing Abstraction Hierarchies with `compare_AH()`

To navigate to other vignettes, go to [Vignette 0 - Welcome to AHgen](Vignette-0---Welcome-to-AHgen.html).

## 1. Geospatial analysis

`AHgen` was originally developed to support analysis of the Urban Systems Abstraction Hierarchy (USAH). A main aim was to create a generic template Abstraction Hierarchy for a UK city, which would be generalisable such that it could be easily modified to reflect specific locations. 

To modify the template USAH in a data-driven manner, the supporting package `OSMtidy` was developed to extract and tidy geospatial data from Open Street Map. `OSMtidy` outputs give a tidy dataset of geospatial data (individual points, lines, and polygons representing different buildings, parks, etc.). Each observation of geospatial data is described by a *desc* term, such as *Attractions; Soft play* and *Sports and recreation; Indoor playground*. In `AHgen`, these *desc* terms are matched to USAH Resources - in this case the Resource *Playgrounds (indoor)*. This is done in an aggregated manner as there are currently 653 *desc* terms and 254 USAH Resources. This process supports an analyst to identify which Resources to filter out from the "master" template version of the USAH, to reflect a specific location.

The `AHgen` functions `check_desc()` and `apply_location()` translate `OSMtidy` outputs into `AHgen` outputs for a specific location.

### 1.1. `check_desc()`

The function `check_desc()` identifies which *desc* terms were present in the Open Street Map data from an `OSMtidy` output, using a key.

This function has two input arguments:

- *geoData* - the `OSMtidy` output, a geospatial dataset
- *key* - the key matching `OSMtidy` *desc* terms to USAH Resources

When applying the function:

- The *geoData* must include a column named *desc*.
- The *key* must include columns named *resource* and *stage_OSMtidy*.

Below is an example application of `check_desc()`.

<details><summary><span style = "color: #2196F3;">Show code - Example 1 - Check `OSMtidy` desc terms to include and exclude</span></summary>
```{r}
# Read in OSMtidy output
geoData = read_rds("../inputs/OSMtidy_Bristol_data_20210826.RDS")

# Read in OSMtidy-AHgen key for allocating OSMtidy desc terms to USAH Resources
key = read_csv("../inputs/OSM-AH-key_3.0_20230717.csv", show_col_types = FALSE)

# Check which desc terms were detected in the Open Street Map data
desc_check = check_desc(geoData = geoData, key = key)

# Inspect a full record of how the key was applied to the OSMtidy output for all 653 desc terms
desc_check$keyApplication
```
Note that if the input geoData is of simple feature class, this can take a while due to reformatting the *geometry* column to character class, so that it can be filtered.
```{r}
# Inspect the USAH Resources that were *not detected* in Open Street Map data 
# but are specified by the key to *always be included*
checkAlwaysInclude = desc_check$notDetected_alwaysInclude
```
Note that these USAH Resources are manually set to "alwaysInclude" in the column *stage_OSMtidy* of the *key* .xlsx file, likely due to being present within a city-scale boundary, but being often-missed or absent in Open Street Map data.

Now you can inspect Resources which would be excluded based on the Open Street Map data alone, and manually check that this is correct before proceeding.
```{r}
# Inspect USAH Resources that were *not detected* in Open Street Map data and *will be excluded* without a manual intervention by the analyst
checkExclude = desc_check$notDetected_exclude
checkExclude %>% select(resource) %>% unique()
```
At this stage the analyst must manually Google each unique USAH Resource & the location name from checkExclude, referring to the shapefile applied to generate the OSMtidy output to ensure any found in the manual searches are inside or outside of the boundary of analysis. This manual check is important to ensure a Resource is not just missing in the Open Street Map data

Then, based on the manual Google check, the analyst must manually create a vector of USAH Resource names that should be included but are currently set to be excluded for this location - as below.
```{r}
# Create vector of USAH Resources to manually include for Bristol
manualInclude_vec_Bristol <- 
  c("Animal shelters", # Multiple
    "Botanical gardens", # University of Bristol Botanic Garden
    "Broadband provider offices", # Virgin Media
    "Diplomatic buildings", # Indian Visa and Consular Centre
    "Electricity provider offices", # Multiple e.g. Bristol Energy
    "Ferry terminals", # Bristol Cruise Terminal
    "Fishing clubs", # Bathampton Angline, Lido, etc.
    "Gas provider offices", # Multiple e.g. Bristol Energy
    "Internet cafes", # At the Well etc.
    "Manufacturing (IT hardware)", # Cubik Innovation performs in-house manufacturing
    "Playgrounds (indoor)", # Jungle Mania etc.
    "Sexual health clinics", # Unity Sexual Health @ Central clinic
    "Springs and wells", # St. Edith's Well
    "Vehicle impounds", # Western Distribution Centre: https://www.bristol.gov.uk/parking/if-your-vehicle-has-been-towed-away
    "Water provider offices") # e.g. Bristol Water

# Create a dataframe of rows 
# to be removed from desc_check$notDetected_exclude
# and then added to desc_check$notDetected_alwaysInclude for this location
manualInclude =
  desc_check$notDetected_exclude %>%
  filter(resource %in% manualInclude_vec_Bristol) %>% ## Specify which manualInclude_vec to use
  mutate(includedBy = "manualCheck")

# Remove from desc_check$notDetected_exclude
desc_check$notDetected_exclude =
  desc_check$notDetected_exclude %>%
  filter(!resource %in% manualInclude_vec_Bristol) ## Specify which manualInclude_vec to use
desc_check$notDetected_exclude

# Add to desc_check$notDetected_alwaysInclude
desc_check$notDetected_alwaysInclude = 
  desc_check$notDetected_alwaysInclude %>%
  mutate(includedBy = "automatedKey") %>%
  rbind(manualInclude)
desc_check$notDetected_alwaysInclude

# Inspect again to ensure this has worked and see the final list of exclusions
checkAlwaysInclude = desc_check$notDetected_alwaysInclude
checkAlwaysInclude %>% select(resource) %>% unique()
checkExclude = desc_check$notDetected_exclude
checkExclude %>% select(resource) %>% unique()
```
</details>


### 1.2. `apply_location()`

The function `apply_location()` applies the *desc_check* to the template baseline version of the Urban Systems Abstraction Hierarchy, to generate a standard `AHgen` output.

It has eight input arguments:

- *desc_check* - the dataset of which USAH Resources to exclude and include for a specific location
- *vInfo_template* - the vertex information for the template baseline USAH
- *edgelist_template* -  the edge list for the template baseline USAH, to be modified for a specific location
- *proxyWeight* - a near-zero weight can sometimes be used to imitate removal of a vertex from the network, without actual removal (e.g. 0.000000001); this is optional and by default this is set to `proxyWeight = 0`
- *AH_name* - the name of the system being modelled by this Abstraction Hierarchy, e.g. "USAH" for Urban Systems Abstraction Hierarchy
- *AH_version* - the version of this Abstraction Hierarchy, e.g. "3.0"; by default this is set to `version = NULL`
- *AH_location* - for applications of the Urban Systems Abstraction Hierarchy to specific geographical locations, e.g. "Bristol"; by default this is set to `location = NULL`
- *AH_scenario* - for applications of the Urban Systems Abstraction Hierarchy to specific scenarios, e.g. "baseline"; by default this is set to `scenario = NULL`

When applying the function:

- *desc_check* must have an element called *notDetected_exclude* with a column named *resource*.
- *vInfo_template* must have a column named *Node*.
- *edgelist_template* must have columns named *layer*, *from*, and *to*.

Below is an example application of `apply_location()`.

<details><summary><span style = "color: #2196F3;">Show code - Example 2 - Generate a standard `AHgen` output for a USAH of a specific location</span></summary>
```{r}
# Read in AHgen output for baseline template USAH
USAH_template_baseline = 
  read_rds("data/template/3.0/USAH_3.0_template_baseline_20230715.RDS")

# Inspect baseline template USAH vInfo
USAH_template_baseline$vInfo

# Inspect baseline template USAH edge list
USAH_template_baseline$edgelist

# Generate baseline Bristol USAH
USAH_Bristol_baseline = 
  apply_location(
    desc_check = desc_check,
    vInfo_template = USAH_template_baseline$vIncluded,
    edgelist_template = USAH_template_baseline$edgelist,
    AH_name = "USAH", 
    AH_version = "3.0", 
    AH_location = "Bristol", 
    AH_scenario = "baseline")

# Inspect the record of the full desc_check
USAH_Bristol_baseline$desc_check

# Inspect the included vertices
USAH_Bristol_baseline$vIncluded

# Inspect the excluded vertices
USAH_Bristol_baseline$vExcluded

# Inspect the adjacency matrix
USAH_Bristol_baseline$adjMat

# Inspect the edge list
USAH_Bristol_baseline$edgelist

# Inspect the igraph
USAH_Bristol_baseline$igraph

# Inspect the results
USAH_Bristol_baseline$results

# Inspect the summary of vertices and edges
USAH_Bristol_baseline$summary
```
</details>


## 2. Scenario analysis
### 2.1. `gen_hazard()`

The function `gen_hazard()` applies an output from the `OSMtidy` function `compare_OSMtidy()`, to generate new weights for USAH edges. 

`compare_OSMtidy()` calculates the proportion of Resources that remain functional after a hazard, based on a baseline count and a scenario count. For example, if a location's baseline count finds that there are 10 of the Resource *Religious buildings*, and a location's scenario count finds that there are 4 of the Resource *Religious buildings* within a 1 in 100-year flood extent, then the capacity of all *Religious buildings* in the area is assumed to be 60% of its baseline capacity. 

Thus the `gen_hazard()` function would generate edges connected the Resource *Religious buildings* with a new weight = 0.6.

The `gen_hazard()` function has five input arguments:

- *edgelist* - the edge list to be modified, usually the edge list for a specific location
- *countCompared* - an `OSMtidy` output comparing the count for a new scenario (e.g. Bristol flood) against a benchmark scenario (e.g. Bristol baseline)
- *vInfo_full* - the full vertex information for all vertices relevant to this Abstraction Hierarchy; in addition to the basic information in vIncluded, it is sometimes useful to include other characteristics e.g. definitions for each vertex, or colour specifications for visualisations; by default this is set to `vInfo_full = NA`
- *hazard* - the name of the hazard, which is used to call a specific column of *vInfo_full* which holds information about which vertices are affected or not affected by a specific hazard; by default this is set to `hazard = NA`
- *proxyWeight* - a near-zero weight can sometimes be used to imitate removal of a vertex from the network, without actual removal (e.g. 0.000000001); this is optional and by default this is set to `proxyWeight = 0`

When applying the function:

- The *edgelist* must contain columns named *to* and *weight*.
- The *countCompared* must contain a list element named *compareBy_physicalObject* which contains columns named *physicalObject* and *total_prop*.
- If used, the *hazard* must be named "flood" or "drought".
- If used, the *vInfo_full* must contain columns named *level*, *subnetwork_hazard_flood*, and *subnetwork_hazard_drought*.

Below is an example application of `gen_hazard()`.

<details><summary><span style = "color: #2196F3;">Show code - Example 3 - Generate new edges representing a hazard</span></summary>
```{r}
# Read in the OSMtidy count for the Bristol baseline
count_Bristol_baseline = 
  read_rds("../inputs/OSMtidy_count_3.0_Bristol_baseline_20210906.RDS") 

# Read in the OSMtidy count for the Bristol flood scenario
count_Bristol_flood = 
  read_rds("../inputs/OSMtidy_count_3.0_Bristol_floodRiver-100_20221212-112538.RDS")

# Compare OSMtidy outputs for a baseline location and flood scenario
# to find the proportion of physical objects that remain functional
countCompared = 
  compareOSMtidy(
    baselineName = "USAH_3.0_Bristol_baseline",
    countOSMtidy_baseline = count_Bristol_baseline,
    scenarioName = "USAH_3.0_Bristol_floodRiver-100",
    countOSMtidy_scenario = count_Bristol_flood)
  
# Generate edgesFlood
edgesFlood = 
  gen_hazard(
    vInfo_template_full = USAH_template_baseline$vInfo,
    edgelist = USAH_Bristol_baseline$edgelist,
    hazard = "flood",
    countCompared = countCompared)
```
In the *edgesFlood* to be introduced, any Resources to be completely removed should also be manually checked with a Google search, as in section 1.1.
```{r}
# Manual checks

# REMOVAL SERVICES TO CHECK IN THE BRISTOL FLOOD EXTENT:
# https://www.google.com/maps/search/bristol+removal+services/@51.4581366,-2.6416926,11.92z

# 29 REMOVAL SERVICES ARE NOT IN THE BRISTOL FLOOD EXTENT:
# Lovely Jubbly Removals, 29 Turtlegate Ave, Bristol BS13 8NN
# DL Removals, 540 Hareclive Rd, Bristol BS13 0LX
# Million Target Ltd, 49, Bristol BS13 9QT
# Langdon Removals Ltd, Block 12, Cater Business Park, Bristol BS13 7TT
# Bristol Man with a Van, 43 Headley Park Ave, Bristol BS13 7NR
# Abbah Removals, 50 Wrington Cres, Bristol BS13 7EP
# JC Man and Van Services, S Liberty Ln, Bristol BS3 2SU
# Bristol Removals, 2a Greville Rd, Southville, Bristol BS3 1LL
# Manley Van Removals, 8 Ilminster Ave, Bristol BS4 1LT
# Rock Steady Removals, 22 Newquay Rd, Bristol BS4 1EA
# KP Transport, 8 Greenwood Rd, Knowle, Bristol BS4 2SX
# Abacus Removals, 821 Bath Rd, Brislington, Bristol BS4 5NL
# Van Move It, 123 Raleigh Rd, Southville, Bristol BS3 1QU
# Kwik Move, Unit 3, Bonville Trading Estate, Bonville Rd, Brislington, Bristol BS4 5QU
# The Little Removals Company, Alpha Rd, Southville, Bristol BS3 1DH
# Excalibur Removals, Ridings Farm, Bridgwater Rd, Dundry, Bristol BS41 8JP
# Bristol Van Removals Ltd, Industrial Park, Unit 3 Netham Rd, Bristol BS5 9PQ
# Smartmove Removals Bristol, Wainbrook House, 1 Hudds Vale Rd, Bristol BS5 7HY
# SWFL Removals and Courier Service, Easton Business Centre, Felix Rd, Easton, Bristol BS5 0HE
# Man with Van Removals Bristol, 34 Cobourg Rd, Montpelier, Bristol BS6 5HX
# Big Van Man Removals, 21 Gordon Ave, Whitehall, Bristol BS5 7DY
# AJM Removals, 569 Fishponds Rd, Bristol BS16 3AF
# Macro Removals, 77 Gloucester Rd, Staple Hill, Bristol BS16 4SL
# Knights of Removals, White Lodge Rd, Bristol BS16 5ND
# Painless Removals, Henleaze House, Henleaze, Bristol BS9 4PN
# Bristol Man and Van, Hambrook Business Park, The Strm, Hambrook, Bristol BS16 1RQ
# We Like 2 Move It, 31 Wrington Cl, Little Stoke, Bristol BS34 6EU
# Wilkinson Removal & Storage, Unit 1b Grange Court Farm Units, Winterbourne, Bristol BS36 1RY
# Get Moved, 34 Mogg St, St Werburgh's, Bristol BS2 9UB

# 3 REMOVAL SERVICES ARE IN THE BRISTOL FLOOD EXTENT:
# Easymove Removals, Albert Cres, Bristol BS2 0SU
# Move On Removals and Storage, 27-29 Baldwin St, Bristol BS1 1LT
# On the Dot Removals Bristol, 21 Horley Rd, St Werburgh's, Bristol BS2 9TL

# THUS 29/32 REMOVAL SERVICES ARE STILL FUNCTIONAL, FOR AN EDGE WEIGHT OF 0.90625

# Manually adjust the weighting of Removal services
edgesFlood = 
  edgesFlood %>%
  mutate(weightNew = ifelse(to == 'Removal services', 0.90625, weightNew))
```
Note that if you would like to export a record of these edges, you can do so with `export_AHgen()`.
```{r}
edgesFlood %>% 
  export_AHgen(
    type = "USAH_edges",
    AH_version = "3.0", 
    AH_location = "Bristol", 
    AH_scenario = "floodRiver-100")
```
</details>


### 2.2. `apply_scenario()`

The function `apply_scenario()` generates a new Abstraction Hierarchy by adjusting an existing Abstraction Hierarchy based on a new edge list. It has seven input arguments:

- *AH_input* - a standard `AHgen` output with eight list elements describing the vertices, network, and results
- *edgelist_scenario* - the new edge list of the Abstraction Hierarchy
- *proxyWeight* - a near-zero weight can sometimes be used to imitate removal of a vertex from the network, without actual removal (e.g. 0.000000001); this is optional and by default this is set to `proxyWeight = 0`
- *AH_name* - the name of the system being modelled by this Abstraction Hierarchy, e.g. "USAH" for Urban Systems Abstraction Hierarchy
- *AH_version* - the version of this Abstraction Hierarchy, e.g. "3.0"; by default this is set to `version = NULL`
- *AH_location* - for applications of the Urban Systems Abstraction Hierarchy to specific geographical locations, e.g. "Bristol"; by default this is set to `location = NULL`
- *AH_scenario* - for applications of the Urban Systems Abstraction Hierarchy to specific scenarios, e.g. "flood"; by default this is set to `scenario = NULL`

When applying the function:

- The *edgelist_scenario* must have used the `weight_edges()` and `weight_hangingVertices()` functions with `remove = FALSE`, or else any edges that are to be excluded from the original Abstraction Hierarchy should have a weight = 0. This enables a record of included and excluded vertices to be noted before removing edges with a weight = 0 and applying network analysis.
- The *edgelist_scenario* must have columns named *layer*, *from*, *to*, and *weight*.
- The *AH_input* must be a list which includes an element named *vIncluded* and an element named *vExcluded*.

<details><summary><span style = "color: #2196F3;">Example 4 - Apply a new edge list to modify an existing Abstraction Hierarchy</span></summary>
```{r}
# Generate new edge list from edgesFlood
edgelist_Bristol_flood = 
  USAH_Bristol_baseline$edgelist %>%
  weight_edges(edgesNew = edgesFlood, remove = FALSE) %>%
  weight_hangingVertices(remove = FALSE)

# Generate a standard AHgen output for the Bristol flood scenario
USAH_Bristol_flood = 
  apply_scenario(    
    AH_input = USAH_Bristol_baseline,
    edgelist_scenario = edgelist_Bristol_flood,
    AH_name = "USAH", 
    AH_version = "3.0", 
    AH_location = "Bristol", 
    AH_scenario = "flood")
```
</details>


## 3. Sensitivity analysis

### 3.1. `calc_sensitivity()`

The function `calc_sensitivity()` finds any edges which are not fully functional (weight = 1) and adjusts their weight by x%. 

It has three input arguments:

- *AH_input*
- *

Note that `calc_sensitivity()` sets any negative values to a minimum threshold of weight = 0 and any positive values > 1 to a maximum threshold of weight = 1.

<details><summary><span style = "color: #2196F3;">Example 5 - Calculate sensitivity analysis</span></summary>
```{r}

```
</details>

### 3.2. `apply_sensitivity()`


<details><summary><span style = "color: #2196F3;">Example 6 - Apply sensitivity analysis to a incorporate in a standard `AHgen` output</span></summary>
```{r}

```
</details>


### 3.3. `summarise_confidence()`


<details><summary><span style = "color: #2196F3;">Example 7 - Summarise confidence of sensitivity analysis results</span></summary>
```{r}

```
</details>


## 4. Comparing Abstraction Hierarchies with `compare_AH()`


## Next
Next in [Vignette 6 - Visualise](Vignette-6---visualise.html), you will see how to visualise Abstraction Hierarchies and network analysis results.

## Last updated
This vignette is associated with `AHgen` v1.0.0, and was last updated by Melissa Bedinger on `r Sys.Date()`.