---
title: "Weight"
author: 
 - Annie Visser-Quinn & Melissa Bedinger
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: paper
    highlight: haddock
vignette: >
  %\VignetteIndexEntry{3 - Weight}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)

rm(list = ls()); cat("\014")

library(knitr)
library(tidyverse)
library(AHgen)
```

<style>
#TOC {
  background: url("images/hex.png");
  background-size: 220px;
  background-position-x: center;
  background-position-y: top;
  padding-top: 280px !important;
  background-repeat: no-repeat;
}
</style>

```{r include = FALSE}
dh = read_adjMat("inputs//mobilityAH_adjMat_20230629.xlsx") 
dv = dh %>% select(level, levelName_full, levelName, Node)
de = dh %>% adjMat_to_edgelist(vInfo = dv)
di = dh %>% adjMat_to_igraph(vInfo = dv)
```
---

In this vignette, you will find out about weighting networks:

- Section 1 - Why weight?
- Section 2 - Weight an edge list with the `tidyverse`
- Section 3 - Weight an igraph with `igraph`
- Section 4 - Weight an edge list with `AHgen`

To navigate to other vignettes, go to [Vignette 0 - Welcome to AHgen](Vignette-0---Welcome-to-AHgen.html).

## 1. Why weight?

A [weighted network](https://en.wikipedia.org/wiki/Weighted_network) is a network where the links between nodes have weights assigned to them. In a number of real-world networks, not all edges in a network are equal. Edges are often assigned weights to represent differences in terms of their strength, intensity, or capacity. For example, a weight could signify the strength of social relationships, the flow of material between within an ecosystem, or the functionality (or loss of functionality) within different parts of an urban system.

## 2. Weight an edge list with the `tidyverse`

In the `tidyverse`, there is a handy `mutate()` function. This adds a new column to a data frame or tibble, of a length equal to the number of rows.

You can check the length of the edge list using `nrow()`. Below is an example for the mobility abstraction hierarchy.

```{r}
de %>% nrow()
```
This means that any new column must be of length 20, i.e. a vector of length 20. Now click on the examples below.

<details><summary><span style = "color: #2196F3;">Show code - Example 1 - Add random weights to an edge list</span></summary>

First, let's use random sampling to create a vector of length 20. This will generate 20 numbers between 0 and 1, which we can use to represent hypothetical weights.

```{r include = FALSE}
set.seed(1)
```
```{r}
vec = sample(x = seq(0, 1, by = 0.01), size = 20)

vec %>% length

vec
```
We can now add this vector to the edge list using `mutate()` to replace the current *weight* column.

```{r}
de_w = de %>% mutate(weight = vec)

de_w
```
</details>

<details><summary><span style = "color: #2196F3;">Show code - Example 2 - Change weights for specific vertices in an edge list</span></summary>

Example 1 used randomly generated weights. In practice we would assign weights with some meaning. 

For example, an ageing vehicle. To reflect this, we could degrade the edges connecting to the vertex *Vehicle components*.

```{r}
de %>%
  mutate(weight = 
           ifelse(from == "Vehicle components" | to == "Vehicle components",
                  0.75, weight)) %>%
  arrange(desc(layer)) 
```

The above code creates a new weight column, where if any row of the edgelist has the term *Vehicle components* in the `to` or `from` columns, it will be assigned a weight of 0.75; for other rows, the weight from before (a weight of 1) will still apply.

This example would require prior coding knowledge to apply. The complexity would increase further if we wanted to change the weightings for multiple edges.

</details>


## 3. Weight an igraph with `igraph`
In [Vignette 2 - Input & Convert](Vignette-2---Input-&-Convert.html), we explored how to extract data and attributes from an igraph. Below is a quick recap.

```{r}
# Extract edges
E(di)

# Extract edge attribute weight
E(di)$weight
```

`E(di)$weight` is effectively an object. Therefore, to change the weighting, we can simply override the weight values using `= vectorOfValues`. 

This vector of values must be the same length as the number of edges. From section 1 we know this is 20, but we can also check it using the igraph function that measures graph size, `gsize()`.

The examples below introduce two ways of weighting the igraph.

<details><summary><span style = "color: #2196F3;">Show code - Example 3 - Add random weights to an igraph</span></summary>

In this example, let's override the weights using `vec`, the vector created in Example 1.

```{r}
E(di)$weight = vec
E(di)$weight
```

It's as simple as that!

</details>

```{r include = FALSE}
di = dh %>% adjMat_to_igraph(vInfo = dv)
```

<details><summary><span style = "color: #2196F3;">Show code - Example 4 - Change weights for specific vertices in an igraph</span></summary>

What if we want to change the weights of specific edges, like those connected to the *Vehicle components* vertex? We need to find out the index, that's the position, of the edge. The mobility abstraction hierarchy is small so this could easily be done by eye.

```{r}
E(di)[18:20]
```

All we'd have to do is overwrite this 'object' with the new weights. 

Let's imagine we don't know the index of the vertex, how can we find out? We begin by finding out the edge names.

```{r} 
edges = E(di) [ from("Vehicle components") ]
edges
```
Now we need to find out the indices of these edges, we can do this using `which()` and `%in%`.

```{r}
index = which(E(di) %in% edges)
index
```

Now we can overwrite the weightings.

```{r}
E(di)$weight[index] = 0.75
E(di)$weight
```

As in Example 2, this requires some prior coding knowledge to apply. It's also more difficult to visualise as the weights and edges can't be inspected side-by-side. The complexity would increase further if we wanted to change the weightings for multiple edges.
</details>

```{r include = FALSE}
di = dh %>% adjMat_to_igraph(vInfo = dv)
```


## 4. Weight an edge list with `AHgen`

Through examples 1-4 we've seen how to change weightings directly. These methods were:

- Cumbersome, requiring prior coding experience
- Impractical for cases where you want to weight edges connected to more than one vertex in a real-world application

`AHgen` has four functions to simplify edge weighting for Abstraction Hierarchies: `check_duplicates()`, `check_diff()`, `weight_edges()`, and `weight_hangingVertices()`.

### 4.1. `check_duplicates()`

The function `check_duplicates()` checks whether a group of edges has any duplicates. If reading in an external .xlsx file, this checks for edges which have been accidentally added twice due to user error. This confirms whether there are any potential conflicts to sort out before introducing the new edge weights, and can also be applied to a finalised edge list to confirm there are no duplicates.

It has one input argument:

- *edges* - the edges to be checked for duplicates

When applying the function:

- the *edges* must include columns named *from* and *to*.

If any conflicts exist, inspect the duplicate edges, and correct the input file as needed before re-loading, re-checking, and proceeding.

<details><summary><span style = "color: #2196F3;">Example 5 - Check new edges for duplicates</span></summary>
```{r}
# Read in new edges
de_w = read_xlsx("inputs//mobilityAH_edges_exampleWeights_20230629.xlsx")
de_w

# Check for duplicates
check_duplicates(edges = de_w)

# See what happens if there are any duplicates
de_w_errorDuplicates = 
  read_xlsx("inputs//mobilityAH_edges_exampleWeights_error-duplicates_20230629.xlsx")
de_w_errorDuplicates
check_duplicates(edges = de_w_errorDuplicates)
```
</details>


### 4.2. `check_diff()`

The `check_diff()` function checks whether new edges to be introduced to an existing edge list has edges which don't match, and if so, returns the non-matching edges. If differences exist, this could introduce an edge which is entirely new to the network rather than simply changing the weight. This function confirms there are no outdated or inconsistent aspects of the network, both in terms of linkages and vertex names.

If any differences exist, inspect these edges, and correct the input file as needed before re-loading, re-checking, and proceeding.

It has two input arguments:

- *edgelist* - the existing edge list
- *edgesNew* - the updated edges to be introduced, to modify the existing edge list

When applying the function:

- Both the *edgelist* and *edgesNew* must include columns named *from* and *to*.

If any conflicts exist, inspect the different edges, and correct the *edgesNew* file as needed before re-loading, re-checking, and proceeding.

<details><summary><span style = "color: #2196F3;">Example 6 - Check differences between two groups of edges</span></summary>
```{r}
# Read in the mobility adjacency matrix
dh = read_adjMat("inputs//mobilityAH_adjMat_20230629.xlsx") 

# Extract vInfo
dv = dh %>% select(level, levelName_full, levelName, Node)

# Convert to edgelist format
de = dh %>% adjMat_to_edgelist(vInfo = dv)

# Read in new edges
de_w = read_xlsx("inputs//mobilityAH_edges_exampleWeights_20230629.xlsx")
de_w

# Check for duplicates
check_diff(edgelist = de, edges = de_w)

# See what happens if there are any differences
de_w_errorDiff = 
  read_xlsx("inputs//mobilityAH_edges_exampleWeights_error-diff_20230629.xlsx")
de_w_errorDiff
check_diff(edgelist = de, edges = de_w_errorDiff)
```
Note that in the second check of *de_w_errorDiff*, there are two differences found. The first is due to a spelling error ("Vehicle capcity" rather than "Vehicle capacity"). The second is due to the introduction of a new Resource which did not exist in the original edge list ("Driver").

</details>


### 4.3. `weight_edges()`

The function `weight_edges()` introduces new weights to an existing edge list. It has three input arguments: 

- *edgelist* - the original edge list to be weighted
- *edgesNew* - the updated edges to be introduced, to modify the existing edge list
- *remove* - an option to keep edges which should be removed (i.e. a weight of 0) in the edge list, or remove them completely; by default this is set to `remove = TRUE` 

When applying the function:

- Both edge lists must have columns named *layer*, *from* and *to*.
- The weight column in the weighted edge list must be called *weightNew*.
- The vertex names must be identical in both edge lists.

The weighted edge list can be generated in R, or created manually in a .csv or .xlsx file, and read into R with functions such as [`read_csv`](https://readr.tidyverse.org/reference/read_delim.html) from the `readr` package or [`read_xlsx`](https://readxl.tidyverse.org/reference/read_excel.html) from the `readxl` package.

Let's explore through some examples.

<details><summary><span style = "color: #2196F3;">Example 7 - Add random weights to an edge list</span></summary>
```{r}
# Function to generate random weights for a specific layer of the Abstraction Hierarchy
gen_exampleWeights <- function(edgelist, layerName = NULL) { 
  
  set.seed(1)
  
  if(!is.null(layerName)) { edgelist <- edgelist %>% filter(layer == layerName) }
  
  output <- 
    edgelist %>%
    rowwise() %>% 
    mutate(weightNew = sample(seq(0.01,1,0.01),1:n())) %>%
    select(-weight)
  
  return(output)
  
}

# Generate random weights for edges in the layer connection levels 4 and 5
de_w = gen_exampleWeights(de, layerName = "l4ORP_l5PO")
de_w

# Introduce the new weights to the edge list
weight_edges(edgelist = de, edgesNew = de_w)
```

It's that easy.

</details>


<details><summary><span style = "color: #2196F3;">Example 8 - Change weights for specific vertices in an edge list</span></summary>


Weightings may be determined externally and imported. Let's revisit our case where a vehicle has been degraded due to age. The *mobilityAH_edges_exampleWeights_20230629.xlsx* spreadsheet has been provided in the *inputs* folder.

```{r}
# Read in edgesNew with new weights
de_w = read_xlsx("inputs//mobilityAH_edges_exampleWeights_20230629.xlsx")
de_w
```

Now we can apply the new weights using `weight_edges()` as before.

```{r}
# Weight the original edge list with edgesNew
weight_edges(edgelist = de, edgesNew = de_w)
```
</details>

### 4.4. `weight_hangingVertices()`

The second function is `weight_hangingVertices()`. When introducing new weights with `weight_edges()`, the `edgesNew` can include edges with a weight of 0, if you intend to completely remove those edges from the original edge list. 

Abstraction Hierarchies are hierarchical networks, wherein if a vertex at level 4 has all of its downward edges to level 5 removed, the vertex at level 4 is no longer required in the network. This means that the vertex at level 4 is now a "hanging" vertex which should also be removed.

The function `weight_hangingVertices()` checks for and removes "hanging" vertices by propagating a check upward layer-by-layer. It has two input arguments: 

- *edgelist* - the edge list to be checked
- *proxyWeight* - a near-zero weight can sometimes be used to imitate removal of a vertex from the network, without actual removal (e.g. 0.000000001); this is optional and by default this is set to `proxyWeight = 0`
- *remove* - an option to keep edges which should be removed (i.e. a weight of 0) in the edge list, or remove them completely; by default this is set to `remove = TRUE` but in some cases it is useful to set `remove = FALSE` e.g. for noting vertices which were removed in a new scenario vs. a benchmark scenario

When applying the function:

- the *edgelist* must have columns named *layer*, *from*, *to*, and *weight*.

<details><summary><span style = "color: #2196F3;">Example 9 - Remove hanging vertices from an edge list</span></summary>
```{r}
# Read in the Urban Systems Abstraction Hierarchy adjacency matrix
dh = read_adjMat("inputs//USAH_3.0_template_baseline_adjMat_20230602.xlsx") 

# Extract vInfo
dv = dh %>% select(level, levelName_full, levelName, Node)

# Convert to edgelist format
de = dh %>% adjMat_to_edgelist(vInfo = dv)

# Read in the edgesNew
de_w = read_xlsx("inputs//USAH_3.0_template_baseline_edges_exampleWeights_20230602.xlsx")

# Inspect the edgesNew
de_w
```
From inspecting the new weights to be introduced, we can surmise that the Resources connected to *Provide dental healthcare services* are degraded, whilst all the Resources connected to *Treat water* should be removed. 

Now we can introduce the new weights, and inspect what would happen if we did not apply the `weight_hangingVertices()` function. Let's try this with the manual `remove = FALSE`, as well as the default `remove = TRUE`, to see how this changes the outputs.
```{r}
# Weight the original edge list with edgesNew
de_removed = weight_edges(edgelist = de, edgesNew = de_w, remove = TRUE)

# Inspect to confirm the level 4 node has been removed
de_removed %>% filter(from == "Treat water" | to == "Treat water")
```

```{r}
# Weight the original edge list with edgesNew
de_zeroed = weight_edges(edgelist = de, edgesNew = de_w, remove = FALSE)

# Inspect to confirm the level 4 node has been removed
de_zeroed %>% filter(from == "Treat water" | to == "Treat water")
```
In both cases we can see that the vertex *Treat water* would still be present in the Urban Systems Abstraction Hierarchy. However, it would have edges only in the upward direction to level 3, and edges in the downward direction to level 5, making it a "hanging" vertex.

To remove it, we must apply the `weight_hangingVertices()` function.
```{r}
# Check for and remove hanging vertices
de_final = weight_hangingVertices(edgelist = de_zeroed, remove = FALSE)

# Inspect to confirm the level 4 node has been removed
de_final %>% filter(from == "Treat water" | to == "Treat water")
```
</details>


After these steps are complete, the new edge list can be converted to other formats with `AHgen`.

### 4.5. A note on using `remove = FALSE`

In some cases it is useful to set `remove = FALSE`, for the purposes of making a list of vertices which should be removed in a new scenario vs. a benchmark scenario. In cases where `weight_edges(remove = FALSE)` and `weight_hangingVertices(remove = FALSE)`, it is important to then remove the edges with a weight = 0 before applying convert functions such as `edgelist_to_igraph()`. 

This is because functions for network analysis - which will be covered in later vignettes - sometimes use the number of vertices in the network as an input to their calculations.

Therefore if you do not remove the edges with a weight = 0 from the final edge list before applying a convert function, any vertices which should be completely removed will still count toward the number of vertices in the network, they will just not be linked to the network - and this will lead to incorrect results.

<details><summary><span style = "color: #2196F3;">Example 10 - Completely remove edges with weight = 0 from an edge list</span></summary>
First inspect the previous example where `remove = FALSE`.
```{r}
# Note that the number of edges in the edge list is 4505
nrow(de_final)

# Note that the number of edges in the edge list with weight = 0 is 7
de_final %>% filter(weight == 0) %>% nrow()
```
Before converting to other data types, you can find out which vertices will be excluded and included, then remove any edges with weight = 0.
```{r}
# Example function to find excluded vertices
find_excluded <- function(edgelist, proxyWeight = 0) {
  
  vertices <- list()
  
  edges_to <- edgelist %>% select(to, weight) %>% rename(Node = to)
  
  edges_from <- edgelist %>% select(from, weight) %>% rename(Node = from)

  edges_all <- edges_to %>% rbind(edges_from) %>% unique()
  
  edges_included <- edges_all %>% filter(weight > proxyWeight) %>% select(-weight)
  
  edges_excluded <- edges_all %>% filter(weight <= proxyWeight) %>% select(-weight)
  
  vertices$excluded <- dplyr::setdiff(edges_excluded, edges_included) %>% pull(Node)
  
  vertices$included <- dplyr::setdiff(edges_included, edges_excluded) %>% pull(Node)
  
  return(vertices)
  
}

# Find the vertices which will be excluded
vertices = find_excluded(edgelist = de_final)

# Inspect the vertices which will be excluded
vertices$excluded

# Then remove the edges with weight = 0
de_final_removed = de_final %>% filter(weight != 0)

# Note that the number of edges is now 4998
nrow(de_final_removed)
```
Now we can convert the updated edge list to other data types!
```{r}
# Update the vInfo for the new network
dv_final_removed = dv %>% filter(!Node %in% all_of(vertices$excluded))

# Then convert to other formats e.g. igraph
de_final_removed %>% edgelist_to_igraph(vInfo = dv_final_removed)
```
Note that at the top of the igraph object, it states there are 482 vertices (1 fewer than the original) and 4498 edges (7 fewer than the original).
</details>


## Next
Next in [Vignette 4 - Analyse - Core functions](Vignette-4---Analyse---Core-functions.html), you will find out about analysing Abstraction Hierarchies with core `AHgen` functions.

## Last updated
This vignette is associated with `AHgen` v1.0.0, and was last updated by Melissa Bedinger on `r Sys.Date()`.