---
title: "1 - Getting started"
author:  
 - Annie Visser-Quinn & Melissa Bedinger
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: paper
    highlight: haddock
vignette: >
  %\VignetteIndexEntry{1 - Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE)

rm(list = ls()); cat("\014")

library(knitr)
library(tidyverse)
library(igraph)
library(AHgen)
```

<style>
#TOC {
  background: url("images/hex.png");
  background-size: 230px;
  background-position-x: center;
  background-position-y: top;
  padding-top: 280px !important;
  background-repeat: no-repeat;
}
</style>

---

In the vignette below, you will find a step-by-step guide to getting started with AHgen.

## Getting started {.tabset}
To get started with AHgen, you'll need to: 

- Step 1 - Install **R** and **RStudio**
- Step 2 - Install the required **R packages**
- Step 3 - Install **AHgen** from a GitHub repository

A walkthrough of each step is provided via the tabs below.


### Step 1 - Install R & RStudio

**What is R?** R is an open source (free) programming language used to develop statistical software such as AHgen. 

**What is RStudio?** RStudio is an "integrated development environment for R". In simpler terms, RStudio provides a more user-friendly environment to run R code from. It is also open source and highly recommended. 

The screenshots below will guide you through the process of downloading [R](https://cloud.r-project.org/) and [RStudio](https://rstudio.com/products/rstudio/download/).

*Note that the versions of R and RStudio shown in the screenshots below will now be outdated; please download the most recent version.*


```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/1 - Getting started/Step 1/1.png"),
               alt = 'install', width = 700)
```
<br>
```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/1 - Getting started/Step 1/2.png"),
               alt = 'install-complete', width = 700)
```
<br>
```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/1 - Getting started/Step 1/3.png"),
               alt = 'install', width = 700)
```
<br>
```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/1 - Getting started/Step 1/4.png"),
               alt = 'install-complete', width = 700)
```
<br>
```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/1 - Getting started/Step 1/5.png"),
               alt = 'install-complete', width = 700)
```
<br>
```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/1 - Getting started/Step 1/6.png"),
               alt = 'install-complete', width = 700)
```
<br>
```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/1 - Getting started/Step 1/7.png"),
               alt = 'install-complete', width = 700)
```
<br>

### Step 2 - R packages

An **R package** is a collection of functions and/or data that have been *packaged* together. For more information, check out [this guide](https://www.datacamp.com/community/tutorials/r-packages-guide). 

#### Required packages to get started 
- [`devtools`](https://cran.r-project.org/web/packages/devtools/readme/README.html) - This package makes package development easier by providing R functions that simplify and expedite common tasks. The install_github() function is used in example AHgen scripts, to install a package straight from GitHub. This is useful if a package is not yet hosted on CRAN, or if you want to play with a different [branch](https://r-bio.github.io/intro-git-rstudio/) of a package (e.g. the newest 'development' version).
- [`pacman`](https://cran.r-project.org/web/packages/pacman/readme/README.html) - This package is an R package management tool that combines the functionality of base library related functions into intuitively named functions. The p_load() function is used in example AHgen scripts, to load multiple packages at the same time instead of using the library() function. Any packages in the p_load() function will also be installed if you have not already done so; this acts as an extra check.

To install these packages:
1. Open RStudio. 
2. Paste `install.packages(c("devtools", "pacman"))` into the console (bottom left hand side).
3. Press `enter`. 
It should look something like below.
*Note, you might be asked to install dependencies or update packages.* 

```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/1 - Getting started/Step 2/1.png"),
               alt = 'install', width = 700)
```
<br>
```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/1 - Getting started/Step 2/2.png"),
               alt = 'install-complete', width = 700)
```
<br>

#### Dependencies

AHgen uses some functions from [`base` R](https://cran.r-project.org/doc/FAQ/R-FAQ.html#R-Add_002dOn-Packages) which do not require you to install any additional packages. In particular, AHgen uses [`base`](https://stat.ethz.ch/R-manual/R-devel/library/base/html/00Index.html), [`grid`](https://stat.ethz.ch/R-manual/R-devel/library/grid/html/00Index.html), [`grDevices`](https://stat.ethz.ch/R-manual/R-devel/library/grDevices/html/00Index.html), and [`stats`](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/00Index.html).

AHgen also relies on functions in several other pre-existing packages. These are called **dependencies**.

The **most important dependency** is the [`tidyverse`](https://www.tidyverse.org/). Instead of one package, the `tidyverse` is actually a collection of R packages designed for data science. All packages share an underlying design philosophy, grammar, and data structures.

**Other dependencies** include: 

*For analysis*

- [**`Biobase`**](https://bioconductor.org/packages/release/bioc/html/Biobase.html) and [`BiocManager`](https://www.bioconductor.org/install/) - The `Biobase` package implements base functions for `Bioconductor`, and has a helpful function called [subListExtract()](https://www.rdocumentation.org/packages/Biobase/versions/2.32.0/topics/subListExtract) to extract the same element from the sublists of a list. For example, if outputs of AHgen are in [list format](https://www.geeksforgeeks.org/r-lists/), and every AHgen output has a list element called 'vertices', the subListExtract() function can extract the list element 'vertices' from multiple lists quickly. This is useful for comparing AHgen outputs. To install `Biobase` you will first need to install the package `BiocManager`. This package serves as the primary way to ensure that the appropriate `Biobase` installation is used with respect to the version of R in use, regardless of the R and `Bioconductor` release cycles.
- [**`data.table`**](https://cran.r-project.org/web/packages/data.table/readme/README.html) provides a high-performance version of base R’s data.frame, with syntax and feature enhancements for ease of use, convenience and programming speed.
- [**`igraph`**](https://cran.r-project.org/web/packages/igraph/index.html) provides routines for simple graphs and network analysis.
- [**`installr`**](https://cran.r-project.org/web/packages/installr/readme/README.html) offers a set of R functions for the installation and updating of software.
- [**`janitor`**](https://cran.r-project.org/web/packages/janitor/readme/README.html) provides simple functions for examining and cleaning dirty data.

*For visualisations*

- [**`ggh4x`**](https://cran.r-project.org/web/packages/ggh4x/readme/README.html),  [**`gginnards`**](https://cran.r-project.org/web/packages/gginnards/readme/README.html), [**`ggraph`**](https://cran.r-project.org/web/packages/ggraph/readme/README.html), [**`ggrepel`**](https://cran.r-project.org/web/packages/ggrepel/readme/README.html), and [**`scales`**](https://cran.r-project.org/web/packages/scales/readme/README.html) are all [`ggplot2`](https://ggplot2.tidyverse.org/) extension packages to customise visualisations.
- [**`plotly`**](https://cran.r-project.org/web/packages/plotly/readme/README.html) helps to create interactive web graphics via the open source JavaScript graphing library [plotly.js](https://github.com/plotly/plotly.js).
- [**`see`**](https://cran.r-project.org/web/packages/see/) provides plotting utilities supporting packages in the [`easystats`](https://cran.r-project.org/web/packages/easystats/index.html) ecosystem  and some extra themes, geoms, and scales for [`ggplot2`](https://ggplot2.tidyverse.org/)
- [**`sf`**](https://cran.r-project.org/web/packages/sf/index.html) provides support for simple features, a standardized way to encode spatial vector data.
- [**`tidytext`**](https://cran.r-project.org/web/packages/tidytext/readme/README.html) provides functions and supporting data sets to allow conversion of text to and from tidy formats, and to switch seamlessly between tidy tools and existing text mining packages.

Some packages are needed to borrow only a single function; other packages are heavily relied upon. For more detail on dependencies you can see `AHgen/data-raw/functionsTracker_20230707.xlsx`.

#### Installation of dependencies

To install BiocManager and Biobase, use the following code.

```{r, eval = FALSE}
# Install BiocManager and Biobase 
# Following instructions from https://www.bioconductor.org/packages/release/bioc/html/Biobase.html

if (!requireNamespace("BiocManager", quietly = TRUE))
install.packages("BiocManager")
BiocManager::install("Biobase")
```

To install the rest of the dependencies, use the following code.

```{r, eval = FALSE}
pacman::p_load(tidyverse, data.table, igraph, installr, janitor, 
               ggh4x, gginnards, ggraph, ggrepel, scales, 
               plotly, see, sf, tidytext)
```
<br>

### Step 3 - Download AHgen

To download the AHgen package itself, use the following code.

```{r, eval = FALSE}
# Use devtools to download the most recent version of AHgen from GitHub
# including any packages that AHgen depends on with "dependencies = TRUE"
devtools::install_github("avisserquinn/AHgen", dependencies = TRUE)
```
<br>

You can load the AHgen package, and see a full list of its current functions, by running the following code.

```{r}
# Load the AHgen package
library(AHgen)

# List the names of AHgen functions and data
ls("package:AHgen")

# List the AHgen function names and required inputs
lsf.str("package:AHgen")
```

### {-}

<br>

## Last updated
This vignette is associated with AHgen v1.0.0, and was last updated by Melissa Bedinger on `r Sys.Date()`.