---
title: "Vignette 5 - Analysis and visualisation"
author: "Dr Annie Visser-Quinn, a.visser-quinn@hw.ac.uk"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    theme: united
vignette: >
  %\VignetteIndexEntry{Vignette 5 - Analysis and visualisation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(
  root.dir = normalizePath('../../AHgen')
  )

rm(list = ls()); cat("\014")

library(knitr)
library(tidyverse)
library(igraph)
```

```{r include = FALSE}
source("functions/functions.R")

dh <- read_adjMat("inputs/AH_mobility.xlsx") 
dv <- dh %>% select(level, levelName, vName)

de <- dh %>% adjMat_to_edgelist

di <- dh %>% adjMat_to_igraph

set.seed(1)
  
de_w <- 
  de %>% 
  rowwise() %>% 
  mutate(weightNew = sample(seq(0.01,1,0.01),1:n())) %>%
  select(-weight)

de_w <- weightEdges(de, de_w)
di_w <- de_w %>% edgelist_to_igraph(dv)
```

## 1. Analysis 
<span style="color: purple;">Explain what type of analysis is done for abstraction hierarchy. Explain centralities (as a minimum).</span>

AHgen has two functions for calculating the unweighted and weighted betweenness centralities, `calcUWVBC()` and `calcWVBC()` respectively. There are two input arguments an igraph and the vertex information. 

```{r message = FALSE, warning = FALSE}
# Calculate unweighted vertex betweenness centrality
di %>% calcUWVBC(dv)

# Weighted edge list
di_w

# Calculate unweighted vertex betweenness centrality
# Note that the outputs are the same
di_w %>% calcUWVBC(dv)

# Calculate weighted vertex betweenness centrality
di_w %>% calcWVBC(dv)
```

<span style="color: purple;">To add these packages to vignette 1</span>

`install.packages(c("tnet", "survival"))`

<span style="color: purple;">Add about how to modify the code</span>

<br>

## 2. Visualisation

```{r}
# key <- read_csv("inputs/visKey.csv", col_types = cols()); key
# 
# # Or create using tribble
# 
# plot_layout <- vis_layout(edgelist = de, 
#                           vInfo = dk, 
#                           minSpacing = 0.05, maxSpacing = 0.25, 
#                           key = key)
# plot_layout
# 
# plot_ggplot <- vis_ggplot(plot_layout, key = NULL)
# plot_ggplot
# 
# plot_plotly <- vis_plotly(plot_ggplot, circles = FALSE)
# plot_plotly
```