---
title: "Application - <br>Introduce flood"
author: 
 - Annie Visser-Quinn & Melissa Bedinger
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: paper
    highlight: haddock
vignette: >
  %\VignetteIndexEntry{Vignette 7C - Application - Introduce flood}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)

rm(list = ls()); cat("\014")

library(knitr)
library(tidyverse)
library(OSMtidy)
library(AHgen)
```

<style>
#TOC {
  background: url("images/hex.png");
  background-size: 220px;
  background-position-x: center;
  background-position-y: top;
  padding-top: 280px !important;
  background-repeat: no-repeat;
}
</style>

---

In this vignette, you will go through the steps of applying a flood extent to a location-specific Urban Systems Abstraction Hierarchy in `AHgen`:

1. Prepare the environment
2. Read in the data
3. Generate new edges to represent the flood hazard
4. Generate Urban Systems Abstraction Hierarchy for Bristol flood scenario
5. Add sensitivity analysis
6. Export

To navigate to other vignettes, go to [Vignette 0 - Welcome to AHgen](Vignette-0---Welcome-to-AHgen.html).


## 1. Prepare the environment

```{r, eval = FALSE}
# Clear the RStudio environment
rm(list = ls()); cat("/014"); gc(verbose = TRUE)

# CTRL + SHIFT + F10 will detach any loaded packages and restart R

# Set the working directory to the folder where this script is saved
setwd(dirname(rstudioapi::getSourceEditorContext()$path)); getwd()

# If you are doing this for the first time

# install the package pacman which checks to see if a package is installed
# and if not it attempts to install the package from CRAN 
# and/or any other repository in the pacman repository list

# and install the package devtools which will allow you to download a package straight from GitHub 
# i.e. if they are not yet on CRAN and/or the pacman repository list

install.packages("devtools")
install.packages("pacman")

# Ensure latest version of AHgen is installed
devtools::install_github("avisserquinn/AHgen", dependencies = TRUE)

# If package has 'non-zero exit status', try re-installing devtools & rlang
install.packages("devtools")
install.packages("rlang")

# Load required packages
pacman::p_load(tidyverse, installr, janitor, OSMtidy, AHgen)
```


## 2. Read in the data

```{r}
# Set name, version, location, baseline, and scenario names
AH_name = "USAH"
AH_version = "3.0"
AH_location = "Bristol"
AH_baseline = "baseline"
AH_scenario = "floodRiver-100"

# Set directory for outputs
directory = "outputs/"

# Read in USAH_template_baseline
USAH_template_baseline = 
  read_rds("inputs/USAH_3.0_template_baseline_20230719-104712.RDS")

USAH_template_baseline

# Read in the USAH_input you want to start with e.g. USAH_location_baseline
USAH_location_baseline = 
  read_rds("inputs/USAH_3.0_Bristol_baseline_20230719-110748.RDS")

# Extract input edgelist
edgelist_location_baseline = USAH_location_baseline$edgelist

# Read in the OSMtidy count for the baseline
count_location_baseline = 
  read_rds("inputs/OSMtidy_count_3.0_Bristol_baseline_20210906.RDS") ## Specify 

# Read in the OSMtidy count for the flood scenario
count_location_flood = 
  read_rds("inputs/OSMtidy_count_3.0_Bristol_floodRiver-100_20221212-112538.RDS")
```


## 3. Generate new edges to represent the flood hazard

```{r}
# Compare OSMtidy outputs for a baseline location and flood scenario
# to find the proportion of physical objects that remain functional
countCompared = 
  compareOSMtidy(
    baselineName = 
      paste0(AH_name, "_", AH_version, "_", AH_location, "_", AH_baseline),
    countOSMtidy_baseline = count_location_baseline,
    scenarioName = 
      paste0(AH_name, "_", AH_version, "_", AH_location, "_", AH_scenario), 
    countOSMtidy_scenario = count_location_flood)
  
# Create edgesFlood
edgesFlood = 
  weight_hazard(
    vInfo_full = USAH_template_baseline$vInfo,
    edgelist = edgelist_location_baseline,
    hazard = "flood", 
    countCompared = countCompared)

# Manual checks

# REMOVAL SERVICES TO CHECK IN THE BRISTOL FLOOD EXTENT:
# https://www.google.com/maps/search/bristol+removal+services/@51.4581366,-2.6416926,11.92z

# 29 REMOVAL SERVICES ARE NOT IN THE BRISTOL FLOOD EXTENT:
# Lovely Jubbly Removals, 29 Turtlegate Ave, Bristol BS13 8NN
# DL Removals, 540 Hareclive Rd, Bristol BS13 0LX
# Million Target Ltd, 49, Bristol BS13 9QT
# Langdon Removals Ltd, Block 12, Cater Business Park, Bristol BS13 7TT
# Bristol Man with a Van, 43 Headley Park Ave, Bristol BS13 7NR
# Abbah Removals, 50 Wrington Cres, Bristol BS13 7EP
# JC Man and Van Services, S Liberty Ln, Bristol BS3 2SU
# Bristol Removals, 2a Greville Rd, Southville, Bristol BS3 1LL
# Manley Van Removals, 8 Ilminster Ave, Bristol BS4 1LT
# Rock Steady Removals, 22 Newquay Rd, Bristol BS4 1EA
# KP Transport, 8 Greenwood Rd, Knowle, Bristol BS4 2SX
# Abacus Removals, 821 Bath Rd, Brislington, Bristol BS4 5NL
# Van Move It, 123 Raleigh Rd, Southville, Bristol BS3 1QU
# Kwik Move, Unit 3, Bonville Trading Estate, Bonville Rd, Brislington, Bristol BS4 5QU
# The Little Removals Company, Alpha Rd, Southville, Bristol BS3 1DH
# Excalibur Removals, Ridings Farm, Bridgwater Rd, Dundry, Bristol BS41 8JP
# Bristol Van Removals Ltd, Industrial Park, Unit 3 Netham Rd, Bristol BS5 9PQ
# Smartmove Removals Bristol, Wainbrook House, 1 Hudds Vale Rd, Bristol BS5 7HY
# SWFL Removals and Courier Service, Easton Business Centre, Felix Rd, Easton, Bristol BS5 0HE
# Man with Van Removals Bristol, 34 Cobourg Rd, Montpelier, Bristol BS6 5HX
# Big Van Man Removals, 21 Gordon Ave, Whitehall, Bristol BS5 7DY
# AJM Removals, 569 Fishponds Rd, Bristol BS16 3AF
# Macro Removals, 77 Gloucester Rd, Staple Hill, Bristol BS16 4SL
# Knights of Removals, White Lodge Rd, Bristol BS16 5ND
# Painless Removals, Henleaze House, Henleaze, Bristol BS9 4PN
# Bristol Man and Van, Hambrook Business Park, The Strm, Hambrook, Bristol BS16 1RQ
# We Like 2 Move It, 31 Wrington Cl, Little Stoke, Bristol BS34 6EU
# Wilkinson Removal & Storage, Unit 1b Grange Court Farm Units, Winterbourne, Bristol BS36 1RY
# Get Moved, 34 Mogg St, St Werburgh's, Bristol BS2 9UB

# 3 REMOVAL SERVICES ARE IN THE BRISTOL FLOOD EXTENT:
# Easymove Removals, Albert Cres, Bristol BS2 0SU
# Move On Removals and Storage, 27-29 Baldwin St, Bristol BS1 1LT
# On the Dot Removals Bristol, 21 Horley Rd, St Werburgh's, Bristol BS2 9TL

# THUS 29/32 REMOVAL SERVICES ARE STILL FUNCTIONAL, FOR AN EDGE WEIGHT OF 0.90625

# Manually adjust the weighting of Removal services
edgesFlood = 
  edgesFlood %>%
  mutate(weightNew = ifelse(to == "Removal services", 0.90625, weightNew))

# Introduce edgesNew to create edgelist_location_flood
# Note that remove = FALSE is required so that we can track vExcluded
# before removing the edges with weight = 0 and converting to an igraph for analysis
edgelist_location_flood = 
  USAH_location_baseline$edgelist %>%
  weight_edges(edgesNew = edgesFlood, remove = FALSE) %>%
  weight_hangingVertices(remove = FALSE)
```


## 4. Generate Urban Systems Abstraction Hierarchy for Bristol flood scenario

```{r}
# Generate output for USAH_scenario
USAH_scenario = 
  apply_scenario(    
    AH_input = USAH_location_baseline,
    edgelist_scenario = edgelist_location_flood,
    AH_name = AH_name, 
    AH_version = AH_version, 
    AH_location = AH_location, 
    AH_scenario = AH_scenario)

# Inspect elements
USAH_scenario$vIncluded
USAH_scenario$vExcluded
USAH_scenario$adjMat
USAH_scenario$edgelist
USAH_scenario$igraph
USAH_scenario$results
USAH_scenario$summary
```


## 5. Add sensitivity analysis

```{r}
# Generate sensitivity analysis
USAH_scenario_sensitivity = 
  apply_sensitivity(
    AH_input = USAH_scenario, 
    AH_name = AH_name, 
    AH_version = AH_version, 
    AH_location = AH_location, 
    AH_scenario = AH_scenario)

# Inspect elements
USAH_scenario_sensitivity[[1]]
USAH_scenario_sensitivity[[2]]
USAH_scenario_sensitivity[[3]]
USAH_scenario_sensitivity$summary$vertices
USAH_scenario_sensitivity$summary$edges
USAH_scenario_sensitivity$vExcluded
USAH_scenario_sensitivity$results
```


## 6. Export

```{r, eval = FALSE}
USAH_scenario_sensitivity %>% 
  export_AHgen(
    type = "USAH",
    directory = directory,
    AH_version = AH_version, 
    AH_location = AH_location, 
    AH_scenario = AH_scenario)

edgesFlood %>% 
  export_AHgen(
    type = "USAH_edges",
    directory = directory,
    AH_version = AH_version, 
    AH_location = AH_location, 
    AH_scenario = AH_scenario)

countCompared %>% 
  export_countCompared(
    directory = directory, 
    version = version, 
    location = location, 
    baseline = baseline, 
    scenario = scenario)
```


## Next
Next in [Vignette 7D - Application - Introduce other hazards](Vignette-7D---Application---Introduce-other-hazards.html), you will go through the steps of applying other hazards to an Urban Systems Abstraction Hierarchy template in `AHgen`.


## Last updated
This vignette is associated with `AHgen` v1.0.0, and was last updated by Melissa Bedinger on `r Sys.Date()`.