---
title: "Application - <br>Apply OSMtidy data"
author: 
 - Annie Visser-Quinn & Melissa Bedinger
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: paper
    highlight: haddock
vignette: >
  %\VignetteIndexEntry{Vignette 7B - Application - Apply OSMtidy data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)

rm(list = ls()); cat("\014")

library(knitr)
library(tidyverse)
library(AHgen)
```

<style>
#TOC {
  background: url("images/hex.png");
  background-size: 220px;
  background-position-x: center;
  background-position-y: top;
  padding-top: 280px !important;
  background-repeat: no-repeat;
}
</style>

---

In this vignette, you will go through the steps of applying `OSMtidy` data to generate a location-specific Urban Systems Abstraction Hierarchy in `AHgen`:

1. Prepare the environment
2. Read in the data
3. Check and edit the Resources to be excluded
4. Generate location-specific Urban Systems Abstraction Hierarchy
5. Export

To navigate to other vignettes, go to [Vignette 0 - Welcome to AHgen](Vignette-0---Welcome-to-AHgen.html).


## 1. Prepare the environment

```{r, eval = FALSE}
# Clear the RStudio environment
rm(list = ls()); cat("/014"); gc(verbose = TRUE)

# CTRL + SHIFT + F10 will detach any loaded packages and restart R

# Set the working directory to the folder where this script is saved
setwd(dirname(rstudioapi::getSourceEditorContext()$path)); getwd()

# If you are doing this for the first time

# install the package pacman which checks to see if a package is installed
# and if not it attempts to install the package from CRAN 
# and/or any other repository in the pacman repository list

# and install the package devtools which will allow you to download a package straight from GitHub 
# i.e. if they are not yet on CRAN and/or the pacman repository list

install.packages("devtools")
install.packages("pacman")

# Ensure latest versions of AHgen is installed
devtools::install_github("avisserquinn/AHgen", dependencies = TRUE)

# If package has 'non-zero exit status', try re-installing devtools & rlang
install.packages("devtools")
install.packages("rlang")

# Load required packages
pacman::p_load(tidyverse, sf, installr, janitor, AHgen)
```


## 2. Read in the data

```{r}
# Set system name, version, location, and scenario
AH_name = "USAH"
AH_version = "3.0"
AH_location = "Bristol"
AH_scenario = "baseline"

# Set directory for outputs
directory = "outputs/"

# Read in USAH_template_baseline
USAH_template_baseline = 
  readRDS("inputs/USAH_3.0_template_baseline_20230719-104712.RDS")

# Read in OSMtidy output for location
geoData = readRDS("inputs/OSMtidy_Bristol_20210826.RDS")

# Read in OSMtidy-AHgen key for allocating OSMtidy descs to AHgen objects
key = read_csv("inputs/OSM-AH-key_3.0_20230717.csv", show_col_types = FALSE)
```


## 3. Check and edit the Resources to be excluded

```{r}
# Check which desc terms were detected in the OSM data
# If the input geoData is of simple feature class, this can take a while 
# due to reformatting geometry to character class which then allows us to filter
desc_check = check_desc(geoData = geoData, key = key)

# Inspect physical objects that were not detected in Open Street Map data 
# but are set by the key to always be included
checkAlwaysInclude = desc_check$notDetected_alwaysInclude

# Inspect physical objects that were not detected in Open Street Map data
# and will be excluded
checkExclude = desc_check$notDetected_exclude

# Google each physical object type + location name 
# to ensure this is not just missing in the Open Street Map data
checkExclude %>% select(resource) %>% unique() 

# Create a vector of Resource names that should be included
# but are currently set to be excluded for this location
manualInclude_vec_Bristol = 
  c("Animal shelters", # Multiple
    "Botanical gardens", # University of Bristol Botanic Garden
    "Broadband provider offices", # Virgin Media
    "Diplomatic buildings", # Indian Visa and Consular Centre
    "Electricity provider offices", # Multiple e.g. Bristol Energy
    "Ferry terminals", # Bristol Cruise Terminal
    "Fishing clubs", # Bathampton Angline, Lido, etc.
    "Gas provider offices", # Multiple e.g. Bristol Energy
    "Internet cafes", # At the Well etc.
    "Manufacturing (IT hardware)", # Cubik Innovation performs in-house manufacturing
    "Playgrounds (indoor)", # Jungle Mania etc.
    "Sexual health clinics", # Unity Sexual Health @ Central clinic
    "Springs and wells", # St. Edith's Well
    "Vehicle impounds", # Western Distribution Centre: https://www.bristol.gov.uk/parking/if-your-vehicle-has-been-towed-away
    "Water provider offices") # e.g. Bristol Water

# Create a dataframe of rows to be removed from desc_check$notDetected_exclude
# and added to desc_check$notDetected_alwaysInclude
manualInclude =
  desc_check$notDetected_exclude %>%
  filter(resource %in% manualInclude_vec_Bristol) %>%
  mutate(includedBy = "manualCheck")

# Remove from desc_check$notDetected_exclude
desc_check$notDetected_exclude =
  desc_check$notDetected_exclude %>%
  filter(!resource %in% manualInclude_vec_Bristol)

desc_check$notDetected_exclude

# Add to desc_check$notDetected_alwaysInclude
desc_check$notDetected_alwaysInclude = 
  desc_check$notDetected_alwaysInclude %>%
  mutate(includedBy = "automatedKey") %>%
  rbind(manualInclude)

desc_check$notDetected_alwaysInclude
```


## 4. Generate location-specific Urban Systems Abstraction Hierarchy

```{r}
# Apply OSMtidy data to USAH
USAH_location = 
  apply_location(
    desc_check = desc_check,
    vInfo_template = USAH_template_baseline$vIncluded,
    edgelist_template = USAH_template_baseline$edgelist,
    AH_name = AH_name, 
    AH_version = AH_version, 
    AH_location = AH_location, 
    AH_scenario = AH_scenario)

# Inspect elements
USAH_location$desc_check
USAH_location$vIncluded
USAH_location$vExcluded
USAH_location$adjMat
USAH_location$edgelist
USAH_location$igraph
USAH_location$results
USAH_location$summary
```


## 5. Export

```{r, eval = FALSE}
# Write output to RDS file
USAH_location %>% 
  export_AHgen(
    type = "USAH",
    directory = directory,
    AH_version = AH_version, 
    AH_location = AH_location, 
    AH_scenario = AH_scenario)
```


## Next
Next in [Vignette 7C - Application - Introduce flood](Vignette-7C---Application---Introduce-flood.html), you will go through the steps of applying a flood extent to a location-specific Urban Systems Abstraction Hierarchy in `AHgen`.


## Last updated
This vignette is associated with `AHgen` v1.0.0, and was last updated by Melissa Bedinger on `r Sys.Date()`.