---
title: "Application - Introduce other hazards"
author: 
 - Annie Visser-Quinn & Melissa Bedinger
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: paper
    highlight: haddock
vignette: >
  %\VignetteIndexEntry{Vignette 7D - Application - Introduce other hazards}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)

rm(list = ls()); cat("\014")

library(knitr)
library(tidyverse)
library(AHgen)
```

<style>
#TOC {
  background: url("images/hex.png");
  background-size: 230px;
  background-position-x: center;
  background-position-y: top;
  padding-top: 280px !important;
  background-repeat: no-repeat;
}
</style>

---

In this vignette, you will go through the steps of applying other hazards to an Urban Systems Abstraction Hierarchy template in `AHgen`:

    0.1. - Prepare the environment
    1.0. - Read in the data
    2.0. - Generate new edges to represent the flood hazard
    3.0. - Generate Urban Systems Abstraction Hierarchy for Bristol flood scenario
    4.0. - Add sensitivity analysis
    5.0. - Export

To navigate to other vignettes, go to [Vignette 0 - Welcome to AHgen](Vignette-0---Welcome-to-AHgen.html).


## 0.1. Prepare the environment

```{r, eval = FALSE}
# Clear the RStudio environment
rm(list = ls()); cat("/014"); gc(verbose = TRUE)

# CTRL + SHIFT + F10 will detach any loaded packages and restart R

# Set the working directory to the folder where this script is saved
setwd(dirname(rstudioapi::getSourceEditorContext()$path)); getwd()

# If you are doing this for the first time
# install the package pacman which checks to see if a package is installed, and if not it attempts to install the package from CRAN and/or any other repository in the pacman repository list
# and install the package devtools which will allow you to download a package straight from GitHub i.e. if they are not yet on CRAN and/or the pacman repository list
# install.packages("devtools")
# install.packages("pacman")

# Ensure latest version of OSMtidy & AHgen are installed
devtools::install_github("avisserquinn/OSMtidy@dev", dependencies = TRUE)
devtools::install_github("avisserquinn/AHgen@dev", dependencies = TRUE)

# If package has 'non-zero exit status', try re-installing devtools & rlang
# install.packages("devtools")
# install.packages("rlang")

# Load required packages
pacman::p_load(tidyverse, installr, janitor, OSMtidy, AHgen)
```


## 1.0. Read in the data

```{r}
# Set name, version, location, baseline, and scenario names
AH_name = "USAH"
AH_version = "3.0"
AH_location = "Bristol"
AH_baseline = "baseline"
AH_scenario = "floodRiver-100"

# Set directory for outputs
directory = "outputs/"

# Read in USAH_template_baseline
USAH_template_baseline = 
  read_rds("inputs/USAH_3.0_template_baseline_20230719-104712.RDS")

USAH_template_baseline

# Read in the USAH_input you want to start with e.g. USAH_location_baseline
USAH_location_baseline = 
  read_rds("inputs/USAH_3.0_Bristol_baseline_20230719-110748.RDS")

# Extract input edgelist
edgelist_location_baseline = USAH_location_baseline$edgelist

# Read in the OSMtidy count for the baseline
count_location_baseline = 
  read_rds("inputs/OSMtidy_count_3.0_Bristol_baseline_20210906.RDS") ## Specify 

# Read in the OSMtidy count for the flood scenario
count_location_flood = 
  read_rds("inputs/OSMtidy_count_3.0_Bristol_floodRiver-100_20221212-112538.RDS")
```


## 2.0. Generate new edges to represent the flood hazard

```{r}
# Compare OSMtidy outputs for a baseline location and flood scenario
# to find the proportion of physical objects that remain functional
countCompared = 
  compareOSMtidy(
    baselineName = 
      paste0(AH_name, "_", AH_version, "_", AH_location, "_", AH_baseline),
    countOSMtidy_baseline = count_location_baseline,
    scenarioName = 
      paste0(AH_name, "_", AH_version, "_", AH_location, "_", AH_scenario), 
    countOSMtidy_scenario = count_location_flood)
  
# Create edgesFlood
edgesFlood = 
  gen_hazard(
    vInfo_full = USAH_template_baseline$vInfo,
    edgelist = edgelist_location_baseline,
    hazard = "flood", 
    countCompared = countCompared)

# Make any manual adjustments
# e.g. Removal centres edge weight in Bristol 1-100 flood scenario

## Bristol 1-100 flood scenario Removal Services to check
## https://www.google.com/maps/search/bristol+removal+services/@51.4581366,-2.6416926,11.92z

## 29 removal services are NOT in 1-100 Bristol flood extent
# Lovely Jubbly Removals, 29 Turtlegate Ave, Bristol BS13 8NN, not in 1-100 flood extent
# DL Removals, 540 Hareclive Rd, Bristol BS13 0LX, not in 1-100 flood extent
# Million Target Ltd, 49, Bristol BS13 9QT, not in 1-100 flood extent
# Langdon Removals Ltd, Block 12, Cater Business Park, Bristol BS13 7TT, not in 1-100 flood extent
# Bristol Man with a Van, 43 Headley Park Ave, Bristol BS13 7NR, not in 1-100 flood extent
# Abbah Removals, 50 Wrington Cres, Bristol BS13 7EP, not in 1-100 flood extent
# JC Man and Van Services, S Liberty Ln, Bristol BS3 2SU, not in 1-100 flood extent
# Bristol Removals, 2a Greville Rd, Southville, Bristol BS3 1LL, not in 1-100 flood extent
# Manley Van Removals, 8 Ilminster Ave, Bristol BS4 1LT, not in 1-100 flood extent
# Rock Steady Removals, 22 Newquay Rd, Bristol BS4 1EA, not in 1-100 flood extent
# KP Transport, 8 Greenwood Rd, Knowle, Bristol BS4 2SX, not in 1-100 flood extent
# Abacus Removals, 821 Bath Rd, Brislington, Bristol BS4 5NL, not in 1-100 flood extent
# Van Move It, 123 Raleigh Rd, Southville, Bristol BS3 1QU, not in 1-100 flood extent
# Kwik Move, Unit 3, Bonville Trading Estate, Bonville Rd, Brislington, Bristol BS4 5QU, not in 1-100 flood extent
# The Little Removals Company, Alpha Rd, Southville, Bristol BS3 1DH, not in 1-100 flood extent (only just)
# Excalibur Removals, Ridings Farm, Bridgwater Rd, Dundry, Bristol BS41 8JP, not in 1-100 flood extent
# Bristol Van Removals Ltd, Industrial Park, Unit 3 Netham Rd, Bristol BS5 9PQ, not in 1-100 flood extent
# Smartmove Removals Bristol, Wainbrook House, 1 Hudds Vale Rd, Bristol BS5 7HY, not in 1-100 flood extent
# SWFL Removals and Courier Service, Easton Business Centre, Felix Rd, Easton, Bristol BS5 0HE, not in 1-100 flood extent
# Man with Van Removals Bristol, 34 Cobourg Rd, Montpelier, Bristol BS6 5HX, not in 1-100 flood extent
# Big Van Man Removals, 21 Gordon Ave, Whitehall, Bristol BS5 7DY, not in 1-100 flood extent
# AJM Removals, 569 Fishponds Rd, Bristol BS16 3AF, not in 1-100 flood extent
# Macro Removals, 77 Gloucester Rd, Staple Hill, Bristol BS16 4SL, not in 1-100 flood extent
# Knights of Removals, White Lodge Rd, Bristol BS16 5ND, not in 1-100 flood extent
# Painless Removals, Henleaze House, Henleaze, Bristol BS9 4PN, not in 1-100 flood extent
# Bristol Man and Van, Hambrook Business Park, The Strm, Hambrook, Bristol BS16 1RQ, not in 1-100 flood extent
# We Like 2 Move It, 31 Wrington Cl, Little Stoke, Bristol BS34 6EU, not in 1-100 flood extent
# Wilkinson Removal & Storage, Unit 1b Grange Court Farm Units, Winterbourne, Bristol BS36 1RY, not in 1-100 flood extent
# Get Moved, 34 Mogg St, St Werburgh's, Bristol BS2 9UB, not in 1-100 flood extent (just)

## 3 removal services ARE in 1-100 Bristol flood extent
# Easymove Removals, Albert Cres, Bristol BS2 0SU, IS in 1-100 flood extent
# Move On Removals and Storage, 27-29 Baldwin St, Bristol BS1 1LT, right on the line - treated as IN 1-100 flood extent
# On the Dot Removals Bristol, 21 Horley Rd, St Werburgh's, Bristol BS2 9TL, IS in 1-100 flood extent

## 29/32 Removal services still functional, for a link weight of 0.90625

# Manual adjustments - Bristol 1-100 flood scenario
edgesFlood = 
  edgesFlood %>%
  mutate(weightNew = ifelse(to == 'Removal services', 0.90625, weightNew))

# Introduce edgesNew to create edgelist_location_flood
# Note that remove = FALSE is required so that we can track vExcluded
# before removing the edges with weight = 0 and converting to an igraph for analysis
edgelist_location_flood = 
  USAH_location_baseline$edgelist %>%
  weight_edges(edgesNew = edgesFlood, remove = FALSE) %>%
  weight_hangingVertices(remove = FALSE)
```


## 3.0. Generate Urban Systems Abstraction Hierarchy for Bristol flood scenario

```{r}
# Generate output for USAH_scenario
USAH_scenario = 
  apply_scenario(    
    AH_input = USAH_location_baseline,
    edgelist_scenario = edgelist_location_flood,
    AH_name = AH_name, 
    AH_version = AH_version, 
    AH_location = AH_location, 
    AH_scenario = AH_scenario)

# Inspect elements
USAH_scenario$vIncluded
USAH_scenario$vExcluded
USAH_scenario$adjMat
USAH_scenario$edgelist
USAH_scenario$igraph
USAH_scenario$results
USAH_scenario$summary
```


## 4.0. Add sensitivity analysis

```{r}
# Generate sensitivity analysis
USAH_scenario_sensitivity = 
  apply_sensitivity(
    AH_input = USAH_scenario, 
    AH_name = AH_name, 
    AH_version = AH_version, 
    AH_location = AH_location, 
    AH_scenario = AH_scenario)

# Inspect elements
USAH_scenario_sensitivity[[1]]
USAH_scenario_sensitivity[[2]]
USAH_scenario_sensitivity[[3]]
USAH_scenario_sensitivity$summary$vertices
USAH_scenario_sensitivity$summary$edges
USAH_scenario_sensitivity$vExcluded
USAH_scenario_sensitivity$results
```


## 5.0. Export

```{r, eval = FALSE}
USAH_scenario_sensitivity %>% 
  export_AHgen(
    type = "USAH",
    directory = directory,
    AH_version = AH_version, 
    AH_location = AH_location, 
    AH_scenario = AH_scenario)
```


## Next
Next in [Vignette 7E - Application - Compare scenarios](Vignette-7E---Application---Compare-scenarios.html), you will go through the steps of comparing Urban Systems Abstraction Hierarchies in `AHgen`.


## Last updated
This vignette is associated with `AHgen` v1.0.0, and was last updated by Melissa Bedinger on `r Sys.Date()`.