---
title: "Application - Generate template"
author: 
 - Annie Visser-Quinn & Melissa Bedinger
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: paper
    highlight: haddock
vignette: >
  %\VignetteIndexEntry{Vignette 7A - Application - Generate template}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)

rm(list = ls()); cat("\014")

library(knitr)
library(tidyverse)
library(AHgen)
```

<style>
#TOC {
  background: url("images/hex.png");
  background-size: 230px;
  background-position-x: center;
  background-position-y: top;
  padding-top: 280px !important;
  background-repeat: no-repeat;
}
</style>

---

In this vignette, you will go through the steps of generating a template Urban Systems Abstraction Hierarchy in `AHgen`:

    0.1. - Prepare the environment
    1.0. - Read in the data
    2.0. - Check the data
    3.0. - Generate the Urban Systems Abstraction Hierarchy template
    4.0. - Export

To navigate to other vignettes, go to [Vignette 0 - Welcome to AHgen](Vignette-0---Welcome-to-AHgen.html).


## 0.1. Prepare the environment

```{r, eval = FALSE}
# Clear the RStudio environment
rm(list = ls()); cat("/014"); gc(verbose = TRUE)

# CTRL + SHIFT + F10 will detach any loaded packages and restart R

# Set the working directory to the folder where this script is saved
setwd(dirname(rstudioapi::getSourceEditorContext()$path)); getwd()

# If you are doing this for the first time
# install the package pacman which checks to see if a package is installed, and if not it attempts to install the package from CRAN and/or any other repository in the pacman repository list
# and install the package devtools which will allow you to download a package straight from GitHub i.e. if they are not yet on CRAN and/or the pacman repository list
# install.packages("devtools")
# install.packages("pacman")

# Ensure latest version of AHgen is installed
devtools::install_github("avisserquinn/AHgen@dev", dependencies = TRUE)

# If package has 'non-zero exit status', try re-installing devtools & rlang
# install.packages("devtools")
# install.packages("rlang")

# Load required packages
pacman::p_load(tidyverse, readxl, installr, janitor, vctrs, AHgen)
```


## 1.0. Read in the data

```{r}
# Set name, version, location, and scenario
AH_name = "USAH"
AH_version = "3.0"
AH_location = "template"
AH_scenario = "baseline"

# Set directory for outputs
directory = "outputs/"

# Read in template USAH adjacency matrix
adjMat_template = 
  read_adjMat("inputs/USAH_3.0_template_baseline_adjMat_20230602.xlsx")

adjMat_template

# Create template USAH vInfo (basic)
vInfo_template = 
  adjMat_template %>% select(level, levelName_full, levelName, Node)

vInfo_template

# Read in template USAH vInfo (full)
vInfo_template_full = 
  read_xlsx("inputs/USAH_3.0_template_baseline_vInfo-full_20230621.xlsx")

vInfo_template_full

# Read in OSMtidy-AHgen key for allocating OSMtidy descs to AHgen objects
key = 
  read_csv("inputs/OSM-AH-key_3.0_20230717.csv", show_col_types = FALSE)

key
```


## 2.0. Check the data

```{r}
# Confirm adjMat has completely symmetrical edges i.e. there are no input errors
adjMat_template %>% check_symmetry()
# If any edges are unsymmetrical check these
# checkSymmetry = adjMat_template %>% check_symmetry()

# Confirm adjMat has no vertices with redundant sets of edges i.e. no vertices
# are connected to the same set of other vertices between two levels
adjMat_template %>% check_redundancy()

# Check node names from different data inputs are compatible
check_names(adjMat_template, vInfo_template_full, key)
# If some node names do not match
# checkNames = check_names(adjMat_template, vInfo_template_full, key)
# checkNames %>% write.csv("checkNames.csv")
```


## 3.0. Generate the Urban Systems Abstraction Hierarchy template

```{r}
# Generate standard AHgen output for USAH baseline template
USAH_template_baseline = 
  gen_AH(vInfo = vInfo_template_full, 
         vIncluded = vInfo_template,
         adjMat = adjMat_template,
         AH_name = AH_name, 
         AH_version = AH_version, 
         AH_location = AH_location, 
         AH_scenario = AH_scenario)

# Inspect elements
USAH_template_baseline$vInfo
USAH_template_baseline$vIncluded
USAH_template_baseline$vExcluded
USAH_template_baseline$adjMat
USAH_template_baseline$edgelist
USAH_template_baseline$igraph
USAH_template_baseline$results
USAH_template_baseline$summary
```


## 4.0. Export

```{r, eval = FALSE}
# Write output to RDS file
USAH_template_baseline %>% 
  export_AHgen(
    type = "USAH",
    directory = directory,
    AH_version = AH_version, 
    AH_location = AH_location, 
    AH_scenario = AH_scenario)

# If subnetwork assignment has been updated, write out to xlsx for future use
# USAH_template_baseline$vInfo %>% 
#   write.xlsx(
#     filenameTimestamp(
#       prefix = paste0(directory,
#                       name, "_", version, "_", location, "_", scenario, 
#                       "_vInfo-full"),
#   extension = ".xlsx"))
```


## Next
Next in [Vignette 7B - Application - Apply OSMtidy data](Vignette-7B---Application---Apply-OSMtidy-data.html), you will go through the steps of applying `OSMtidy` data to generate a location-specific Urban Systems Abstraction Hierarchy in `AHgen`.


## Last updated
This vignette is associated with `AHgen` v1.0.0, and was last updated by Melissa Bedinger on `r Sys.Date()`.