---
title: "Analyse - Core functions"
author: 
 - Annie Visser-Quinn & Melissa Bedinger
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: paper
    highlight: haddock
vignette: >
  %\VignetteIndexEntry{4 - Analyse - Core functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)

rm(list = ls()); cat("\014")

library(knitr)
library(tidyverse)
library(igraph)
library(AHgen)
```

<style>
#TOC {
  background: url("images/hex.png");
  background-size: 230px;
  background-position-x: center;
  background-position-y: top;
  padding-top: 280px !important;
  background-repeat: no-repeat;
}
</style>
---

In this vignette, you will find out about analysing Abstraction Hierarchies:

- Section 1 - Network metrics
- Section 2 - `calc_degrees()`
- Section 3 - `calc_EC()`
- Section 4 - `calc_SBC()`
- Section 5 - `gen_results()`
- Section 6 - `summarise_AH()`


## 1. Network metrics

Network analysis is a useful tool to quantify and analyse the interactions and structure of graphs, which represent the relationships among entities such as sectors within an urban system. For more background, check out [this guide by Barabasi](http://networksciencebook.com/).

There are numerous metrics which can be calculated to analyse networks. Network metrics help to describe a network at either a local level (which looks at individual elements i.e. particular nodes or edges) or a global level (which looks at the network as a whole). The current verison of `AHgen` focuses on analysing the network at a local level, using the metrics [Degree](https://en.wikipedia.org/wiki/Degree_(graph_theory)), [Eigenvector Centrality (EC)](https://en.wikipedia.org/wiki/Eigenvector_centrality), and [Stable Betweenness Centrality (SBC)](https://doi.org/10.1109/ICASSP.2014.6854324). For more detail, and a longer list of options for network analysis metrics, check out [this review of network metrics by Morrison et al.](https://doi.org/10.1007/s41109-022-00476-w).


## 2. `calc_degrees()`

The (node) degree is simply the number of edges that are connected to a vertex. In a hierarchical network like an Abstraction Hierarchy, multiple types of degree can be calculated. For each vertex, the following are calculated:

- the *up* degree - the number of edges to the level above the vertex
- the *down* degree - the number of edges to the level below the vertex
- the *total* degree - the sum of the *up* and *down* degrees

Because an Abstraction Hierarchy has five levels, the *up* and *down* degrees are also calculated for "additional" steps away from the vertex. 

For example, for a Resource vertex at level 5, we first calculated the *up* degree to level 4 (*nodeDegree_up1*) which is equal to the *nodeDegree_total* due to the vertex being at the bottom of the hierarchy and having no edges downward. We can then propagate upwards to calculate:

- for the connected vertices found at level 4, the *up* degree to vertices in level 3 (*nodeDegree_up2*)
- for the connected vertices found at level 3, the *up* degree to vertices in level 2 (*nodeDegree_up3*), and
- for the connected vertices found at level 2, the *up* degree to vertices in level 1 (*nodeDegree_up4*)

Another example would be a Task vertex at level 3. After calculating the *up* degree to level 2 (*nodeDegree_up1*) and the *down* degree to level 4 (*nodeDegree_down1*), we can then propagate in either direction to also calculate:

- for the connected vertices found at level 2, the *up* degree to vertices in level 1 (*nodeDegree_up2*)
- for the connected vertices found at level 4, the *down* degree to vertices in level 5 (*nodeDegree_down2*)

This is done with the function `calc_degrees()`. It has two input arguments: 

- *igraph* - the igraph to be analysed
- *vInfo* - the vertex information

When applying the function:

- The vertex names must be identical in both the igraph and the vInfo

Below is an example of the application of `calc_degrees()`.

<details><summary><span style = "color: #2196F3;">Example 1 - Degree(s)</span></summary>
```{r}
# Read in adjacency matrix for the Urban Systems Abstraction Hierarchy
dh = read_adjMat("../inputs/USAH_3.0_template_baseline_adjMat_20230602.xlsx") 

# Extract vInfo
dv = dh %>% select(level, levelName_full, levelName, Node)

# Convert to an igraph
di = adjMat_to_igraph(adjMat = dh, vInfo = dv)

# Calculate the degree(s) of each vertex
degrees = calc_degrees(igraph = di, vInfo = dv)
```
</details>


## 3. `calc_EC()`
<details><summary><span style = "color: #2196F3;">Example 1 - Degree(s)</span></summary>
```{r}
```
</details>


## 4. `calc_SBC()`
<details><summary><span style = "color: #2196F3;">Example 1 - Degree(s)</span></summary>
```{r}
```
</details>


## 5. `gen_results()`
<details><summary><span style = "color: #2196F3;">Example 1 - Degree(s)</span></summary>
```{r}
```
</details>


## 6. `summarise_AH()`
<details><summary><span style = "color: #2196F3;">Example 1 - Degree(s)</span></summary>
```{r}
```
</details>


## Last updated
This vignette is associated with `AHgen` v1.0.0, and was last updated by Melissa Bedinger on `r Sys.Date()`.