---
title: "Welcome to AHgen"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: paper
    highlight: haddock
vignette: >
  %\VignetteIndexEntry{Vignette 1 - Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(
  root.dir = normalizePath('../../AHgen')
  )

rm(list = ls()); cat("\014")

library(knitr)
library(tidyverse)
library(igraph)
```

<span style = "color:red;">@annie This document will be the readme. Add preamble text.</span>

## Abstraction hierarchy 

<span style="color:red">@melissa Couple of lines for introduction. Describe various things (text might not go exactly here): AH; AHgen. **AHgen** is provided as software code (i.e. it has not been packaged); vertex; layer; level; weighting; flow chart(s); effectively point user in right direction with some general background. </span>


## Getting started {.tabset}
To get started, you'll need to: 

- Step 1 - **Install** R and (optional) RStudio
- Step 2 - Install the required R **packages**
- Step 3 - **Download the AHgen** source code from this GitHub repository
- Step 4 - Prepare a script in R to **load the AHgen functions**

A walkthrough of each step is provided via the **tabs** below.


### Step 1 - Install

These screenshots will guide you through the process of downloading [R](https://cloud.r-project.org/) and [RStudio](https://rstudio.com/products/rstudio/download/). But first:

- **What is R?** - R is an open source (free) programming language used to develop statistical software such as AHgen. 
- **What is RStudio?** - RStudio is an "integrated development environment for R". In simpler terms, RStudio provides a more user-friendly environment to run R code from. It is also open source and highly recommended. 

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.align = "center"}
files <- list.files("vignettes/images/gettingStarted", pattern = ".png", full.names = TRUE)

include_graphics(normalizePath(files))
```

### Step 2 - R packages

**Required packages:** `readxl`, `openxlsx`, `tidyverse`, `igraph`, `tnet` and `ggraph`

An **R package** is a collection of functions and/or data that have been *packaged* together. For more information, check out [this handy guide](https://www.datacamp.com/community/tutorials/r-packages-guide). 

To install the required packages:

- Open RStudio
- Paste `install.packages(c("readxl", "openxlsx", "tidyverse", "igraph", "tnet", "ggraph"))` into the console (bottom left hand side) and press `enter`. *Note, you might be asked to install dependencies or update packages*

A package is like a book, you use a `library()` to check a package out of the library. AHgen will load these packages for you, or you can copy and paste the code chunk below.

```{r, eval = FALSE}
library(readxl)
library(openxlsx)
library(tidyverse)
library(igraph)
library(tnet)
library(ggraph)
```

### Step 3 - Download AHgen
<span style="color: red;">To be filled in once the github is more up to date.</span>

### Step 4 - Load AHgen functions
The functions that make up AHgen are located in the *functions* folder. You can load all of them in by running the script in *functions.R* using `source()`. Step-by-step: 

- Open RStudio
- Navigate to *File > New file > R Script*
- Save the script in the AHgen directory (folder)
- Copy and paste the code chunk below into your script

```{r, eval = FALSE}
rm(list = ls()) # Clear environment
cat("\014") # Clear console

setwd(dirname(rstudioapi::getSourceEditorContext()$path)) # Set working directory to script directory

source("functions/functions.R") # Read in the functions script
```

- Highlight your script and run by pressing `Ctrl+Enter`. The names of the functions will appear in the top right corner
- Type `ls()` directly in the console to print the names of the loaded functions, shown below

```{r, include = FALSE}
source("functions/functions.R")
```
```{r, echo = FALSE}
ls()
```

### {-}


## Vignettes
Vignettes are the documentation that show you how to use AHgen. They are provided in the **vignettes** folder of AHgen in a **.html** file which can be read in your browser. 

Two types of vignette are provided:

- **Vignettes <span style = "color:red;">X-Y</span>** - Designing to get you on your way, introducing the AHgen functions through practical examples
- **Vignettes <span style = "color:red;">A-B</span>** - AHgen workflows for three example cases