---
title: "Input & Convert"
author: 
 - Annie Visser-Quinn & Melissa Bedinger
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: paper
    highlight: haddock
vignette: >
  %\VignetteIndexEntry{2 - Input & Convert}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)

rm(list = ls()); cat("\014")

library(knitr)
library(tidyverse)
library(AHgen)
```

<style>
#TOC {
  background: url("images/hex.png");
  background-size: 230px;
  background-position-x: center;
  background-position-y: top;
  padding-top: 280px !important;
  background-repeat: no-repeat;
}
</style>

---

In this vignette, you will find out about inputs and data in `AHgen`:

- Section 1 - Data types
- Section 2 - Read in data
- Section 3 - Check the data
- Section 4 - Convert to different data types
- Section 5 - Navigate the data using the `tidyverse` and `igraph` packages

For other vignettes, go to [Vignette 0 - Welcome to AHgen](Vignette-0---Welcome-to-AHgen.html).

## 1. Data types

### 1.1 What is an adjacency matrix?

An [adjacency matrix](https://proofwiki.org/wiki/Definition:Adjacency_Matrix) is a matrix which describes a network by representing which vertices are adjacent (i.e. connected) to which other vertices. If there are N vertices in a network, there are N columns (one for each vertex) and N rows (one for each vertex) in the matrix. Each cell of the matrix includes number signifying the connection between the two vertices, representing a potential edge or link.

An adjacency matrix for a simple graph has elements which are all either 0 (representing no connection between the two vertices) or 1 (representing a connection between the two vertices). This is a binary representation of a network. An adjacency matrix for a weighted network contains the weights of the edges (e.g. a weight of 0.75). A higher weight represents a stronger connection than a lower weight (e.g. an edge with a weight of 0.75 is stronger than an edge with a weight of 0.25).

Abstraction Hierarchies are treated as undirected graphs, thus an adjacency matrix for an Abstraction Hierarchy is symmetrical about the main diagonal.

*An example adjacency matrix*
```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/2 - Input & Convert/mobility_adjMat.png"), alt = 'Adjacency Matrix XLSX', width = 800)
```

<br>

### 1.2 What is an edgelist? 

An [edge list](https://en.wikipedia.org/wiki/Edge_list) is a data structure which describes a network by listing each connection as a row of information. The first column describe the vertex the edge is coming 'from', the second column describes the vertex the edge is coming 'to', and if it is a weighted network the third column gives the numeric value of the 'weight'. In an undirected network such as the Abstraction Hierarchy, the 'from' and 'to' columns are merely representative of an undirected edge between two vertices rather than literally being 'from' and 'to' one vertex or another.

*An example edge list*
```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/2 - Input & Convert/mobility_edges.png"), alt = 'Edges XLSX', width = 500)
```

<br>

### 1.3 What is an igraph?

An igraph is an object class from the package of the same name. The igraph data type is a special way of formatting complex networks, enabling the creation of large networks and speedy application of network analysis. 

The code output below is an *igraph of an abstraction hierarchy*.

```{r echo = FALSE}
dh = read_adjMat("../inputs/microwaveAH_adjMat_20230629.xlsx") 
dv = dh %>% select(level, levelName_full, levelName, Node)
di = dh %>% adjMat_to_igraph(vInfo = dv)
di
```
The first line states that this is an igraph. It also provides a codename for the graph, *UNW-*:

- **U** The graph is **U**ndirected
- **N** The vertices have **N**ames
- **W** The **W**eight attribute has been set
- **-** The graph is not bipartite

The second line provides the attributes of the **vertices** (denoted as **v**) and **edges** (denoted as **e**):

- **name (v/c)** - The vertices are named, and have the class **c**, character.
- **level (v/n)** - The vertices have an attribute called **level**, with the class **n**, numeric. This denotes which level (from 1 through 5) a vertex belongs to.
- **levelName_full (v/c)** - The vertices have an attribute called **levelName_full**, with the class **c**, character. This gives the full level name in human factors and Cognitive Work Analysis terminology (from 'Functional purposes' through 'Physical objects').
- **levelName (v/c)** - The vertices have an attribute called **levelName**, with the class **c**, character. This gives a more concise level name (from 'Purposes' through 'Resources').
- **weight (e/n)** - The edges have a weight, with the class **n**, numeric.
- **layer (e/c)** - The edges have an attribute called **layer**, with the class **c**, character.

The remainder of the output represents the edges and the vertex names. 


## 2. Read in data

Here we focus on the two functions, `read_adjMat()` and `read_vInfo()`, which read in the main inputs to `AHgen`:

- An **adjacency matrix** capturing the edges between the levels of the abstraction hierarchy.
- **Vertex information** describing the level and level name associated with each vertex.
These inputs should be in the .xlsx format, i.e. an Excel spreadsheet.


### 2.1 Adjacency matrix

The adjacency matrix can be read in using the function `read_adjMat()`. This function has three input arguments:

- `filename` - The file path (relative to the `AHgen` folder).
- `sheet` - The sheet number of the adjacency matrix; this is set to `sheet = 1` by default.
- `rescale` - Whether the edge weightings in the adjacency matrix should be rescaled to 0.5. This is sometimes useful for testing the hypothetical degradation of an Abstraction Hierarchy network. This is set to `rescale = FALSE` by default.
*Note, two of the three `read_adjMat()` input arguments have default values. If only the filename is provided, these arguments will take this default value.*

Click below to explore two example applications.

<details><summary><span style = "color: #2196F3;">Show code - Example 1 - Read adjacency matrix input (microwave)</span></summary>
The classic human factors example of cooking food in a microwave. The Excel file, *microwaveAH_adjMat_20230629.xlsx* is provided in the *inputs* folder. An annotated exert of the adjacency matrix is shown below.

1. The column of vertex names must be named **Node**.
2. Vertex names run left to right.
3. And top to bottom.
4. The **edges** between vertices are denoted by cells with a value of **1**. **Non-edges** are denoted by **empty** cells. *Note, zeroes can also be used to indicate non-edges*.
5. The diagonal cannot be linked i.e. a vertex cannot link to itself.

```{r }
read_adjMat("../inputs/microwaveAH_adjMat_20230629.xlsx")
```

Notice that the empty cells from the Excel file are filled with zeroes by `AHgen`. This means these vertices are not linked, i.e. they have no weight. We might want to rescale the linked, or weighted edges, to 0.5. To do this, change the `rescale` argument to `TRUE`.

```{r}
read_adjMat("../inputs/microwaveAH_adjMat_20230629.xlsx", rescale = TRUE)
```
</details>


<details><summary><span style = "color: #2196F3;">Show code - Example 2 - Read adjacency matrix input (mobility)</span></summary>

We now consider a second example abstraction hierarchy, *mobilityAH_adjMat_20230629.xlsx*, which is also available in the *inputs* folder.

1. As before, the column of vertex names is called **Node**.
2. And the vertex names run left to right.
3. And top to bottom.
4. The **edges** between vertices are again denoted by cells with a value of **1**. This time, **non-edges** are denoted by **0** cells.
5. But there are three extra columns of vertex information, **level**, **levelName_full**, and **levelName**. If these columns had different names they would be read in as vertices!

With only one sheet, the first sheet is the adjacency matrix, and if we don't want to rescale the data, then only one input argument is required `"../inputs/mobilityAH_adjMat_20230629.xlsx"`.

```{r}
read_adjMat("../inputs/mobilityAH_adjMat_20230629.xlsx")
```

The code chunk below shows what happens if there is a mismatch in column names. Check out the file *mobilityAH_error_adjMat_20230629.xlsx* to see if you can spot the problem.

```{r, error = TRUE}
read_adjMat("../inputs/mobilityAH_error_adjMat_20230629.xlsx")
```
</details>


### 2.2 Vertex information
The vertex information can be read in using the function `read_vInfo()`. This function has two input arguments:

- `filename` - The file path (relative to the `AHgen` folder).
- `sheet` - The sheet number of the vertex information. This is set to `sheet = 1` by default.

<details><summary><span style = "color: #2196F3;">Show code - Example 3 - Read vInfo input (microwave)</span></summary>

For the microwave AH, the vertex information is in the second sheet. Each row describes a single vertex in terms of level, level name (full), level name, and vertex name. To avoid errors, these columns must be named **level**, **levelName_full**, **levelName** and **Node**. 

```{r}
read_vInfo("../inputs/microwaveAH_vInfo_20230629.xlsx")
```
</details>


<details><summary><span style = "color: #2196F3;">Show code - Example 4 - Read vInfo input (mobility)</span></summary>

We saw in Example 2 that the vertex information was embedded in the adjacency matrix. This can be extracted using the `tidyverse` function `select`.

```{r}
read_adjMat("../inputs/mobilityAH_adjMat_20230629.xlsx") %>% 
  select(level, levelName_full, levelName, Node)
```
</details>


## 3. Check the data

There are several `AHgen` functions to perform checks on data and outputs. Here we focus on the functions to check raw data after it has been read in from a spreadsheet format. These are `check_symmetry`, `check_redundancy`, and `check_names`.

<details><summary><span style = "color: #2196F3;">Show code - Example 5 - Check symmetry</span></summary>

The `check_symmetry` function confirms that an adjacency matrix has completely symmetrical edges. This is most important for an undirected graph where an adjacency matrix must be symmetrical along its diagonal. If the adjacency matrix is not symmetrical, `check_symmetry` will return a data frame of inconsistent edges to check.
```{r}
# Confirm adjMat has completely symmetrical edges i.e. there are no input errors
dh_USAH = 
  read_adjMat("../inputs/USAH_3.0_template_baseline_adjMat_20230602.xlsx")

symmetryCheck = dh_USAH %>% check_symmetry()

# See what happens when a few of the edges in the adjacency matrix are unsymmetrical
dh_USAH_error = 
  read_adjMat(
  "../inputs/USAH_3.0_template_baseline_error-symmetry_adjMat_20230602.xlsx")

symmetryCheck_error = dh_USAH_error %>% check_symmetry()
symmetryCheck_error
```
</details>


<details><summary><span style = "color: #2196F3;">Show code - Example 6 - Check redundancy</span></summary>

The `check_redundancy` function confirms that an adjacency matrix has no vertices which with redundant sets of edges. In the development of the Urban Systems Abstraction Hierarchy, a guiding principle of development was that if two Resources were connected to the same set of Processes, the two Resources should be combined. For example, if *Coworking office spaces* and *Offices (other)* fulfilled the same Processes, these were combined into an umbrella Resource. If the adjacency matrix contains vertices with redundant sets of edges, `check_redundancy` will return a dataframe of these vertices to check.
```{r}
# Confirm adjMat has no vertices with redundant sets of edges i.e. no vertices
# are connected to the same set of other vertices between two levels
dh_USAH = 
  read_adjMat("../inputs/USAH_3.0_template_baseline_adjMat_20230602.xlsx")

redundancyCheck = dh_USAH %>% check_redundancy()

# See what happens when a few of the edges in the adjacency matrix are unsymmetrical
dh_USAH_error = 
  read_adjMat(
    "../inputs/USAH_3.0_template_baseline_error-redundancy_adjMat_20230602.xlsx")

redundancyCheck_error = dh_USAH_error %>% check_redundancy()
```
Note that in the correct Urban Systems Abstraction Hierarchy, the difference between *Coworking office spaces* and *Offices (other)* are that:
- *Coworking office spaces* fulfils the Process *Provides other discretionary services* while *Offices (other)* does not
- *Offices (other)* fulfils the Process *Provides IT services* while *Coworking office spaces* does not
</details>


<details><summary><span style = "color: #2196F3;">Show code - Example 7 - Check names</span></summary>

In applications of `AHgen` to the Urban Systems Abstraction Hierarchy, outputs from another software `OSMtidy` are sometimes used. `OSMtidy` outputs give a tidy dataset of geospatial data (individual points, lines, and polygons representing different buildings, parks, etc.). Each observation of geospatial data is described by a *desc* term, such as *Attractions; Soft play* and *Sports and recreation; Indoor playground*. In `AHgen`, these *desc* terms are matched to Abstraction Hierarchy Resources in an aggregated manner - in this case the Resource *Playgrounds (indoor)*.

The `check_names` function confirms that *desc* name and the *Resource* name are both matched in a key. If there are any unmatched names (an additional *desc* term in the `OSMtidy` output, or an additional vertex in the Abstraction Hierarchy), `check_names` will return a dataframe of these to check. Usually this means that a new *desc* term or changed Resource name needs to be added to the key; in some cases an adjustment to the USAH has been made but a spelling error exists, and this check catches it before any issues arise in further `AHgen` analysis.
```{r}
# Read in adjacency matrix
dh_USAH = 
  read_adjMat("../inputs/USAH_3.0_template_baseline_adjMat_20230602.xlsx")

# Read in vInfo
dv_USAH = 
  read_vInfo("../inputs/USAH_3.0_template_baseline_vInfo-full_20230621.xlsx")

# Read in key to match names from one software package (OSMtidy) 
# to names in another software package (AHgen)
key = read_csv(
  "../inputs/OSM-AH-key_3.0_20210906.csv", show_col_types = FALSE) 

# Check that all the names match and return any that do not match
nameCheck_correct = 
  check_names(adjMat_template = dh_USAH, vInfo_template = dv_USAH, key = key)

# See what happens when there is a non-matching name
# Read in adjacency matrix
dh_USAH_error = 
  read_adjMat(
    "../inputs/USAH_3.0_template_baseline_error-names_adjMat_20230602.xlsx")

# Read in vInfo
dv_USAH_error = 
  read_vInfo(
    "../inputs/USAH_3.0_template_baseline_error-names_vInfo-full_20230621.xlsx")

# Read in key to match names from one software package (OSMtidy) 
# to names in another software package (AHgen)
key = read_csv(
  "../inputs/OSM-AH-key_3.0_20210906.csv", show_col_types = FALSE)

# Check that all the names match and return any that do not match
nameCheck_error = 
  check_names(
    adjMat_template = dh_USAH_error, vInfo_template = dv_USAH_error, key = key)
```
The object `nameCheck_error` lets us know that there is one item present in the Abstraction Hierarchy Resources level, that is not present in the key: *Tardis*.
</details>


## 4. Convert to different data types
In `AHgen`, the abstraction hierarchy can be captured in three different formats:

- Adjacency matrix
- Edge list
- igraph

To change format, `AHgen` has a family of six conversion functions:

- `adjMat_to_edgelist()` and `adjMat_to_igraph()`, where the input arguments are the adjacency matrix
- `edgelist_to_igraph()` and `edgelist_to_adjMat()`, where the input arguments are the abstraction hierarchy in edge list format, and the vertex information
- `igraph_to_adjMat()` and `igraph_to_edgelist()`, where the input arguments are the abstraction hierarchy in the igraph format

Before introducing the examples, we need to save the adjacency matrix and vertex information as objects in the environment. We do this using the `<-` or `=` operators.

```{r}
dh = read_adjMat("../inputs/mobilityAH_adjMat_20230629.xlsx") 
dv = dh %>% select(level, levelName_full, levelName, Node)
```

<details><summary><span style = "color: #2196F3;">Show code - Example 8 - Convert from adjacency matrix</span></summary>
```{r}
de = dh %>% adjMat_to_edgelist(vInfo = dv); de
di = dh %>% adjMat_to_igraph(vInfo = dv); di
```
</details>

<details><summary><span style = "color: #2196F3;">Show code - Example 9 - Convert from edge list</span></summary>
Note that, when converting from an edge list, we need to specify the vertex information. 

```{r}
de %>% edgelist_to_igraph(vInfo = dv)
de %>% edgelist_to_adjMat(vInfo = dv)
```
</details>

<details><summary><span style = "color: #2196F3;">Show code - Example 10 - Convert from igraph</span></summary>
```{r}
di %>% igraph_to_adjMat()
di %>% igraph_to_edgelist()
```

<br>

## 5. Navigate the data

Now let's consider how to navigate and explore these three data types.

- The **adjacency matrix** and **edge list** are in a tibble format, an easy to read tabular form. They can be navigated like any data.frame or tibble in R using functions such as `select()` and `filter()`.
- The **igraph** is a special way of formatting complex networks. The package of the same name introduces a number of functions to interrogate the network.

Click below for examples for each data type. 

<details><summary><span style = "color: #2196F3;">Show code - Example 11 - Inspect adjacency matrix</span></summary>
```{r}
# Select vertices by name
dh %>% select(`Provide efficient and accessible mobility`)

# Select vertices which match a pattern
dh %>% select(Node, contains("safe"))

# Filter to a specific vertex
dh %>% filter(Node == "Compliant user behaviour")

# Filter to vertices matching a pattern
dh %>% filter(str_detect(Node, "capab|Capab"))
```
</details>

<details><summary><span style = "color: #2196F3;">Show code - Example 12 - Inspect edge list</span></summary>

```{r}
# Filter for edges in a specific layer
de %>% filter(layer == "l1FP_l2VPM")

# Filter for edges linking vertices containing road
de %>% filter_all(any_vars(str_detect(str_to_lower(.), "road")))

# Filter for edges linking from the vertex "Vehicle capacity"
de %>% filter(from == "Vehicle capacity")
```
</details>

<details><summary><span style = "color: #2196F3;">Show code - Example 13 - Inspect igraph</span></summary>
```{r}
# Extract vertices
V(di)

# Extract edges
E(di)

# Extract vertex attribute name
V(di)$name

# Extract vertex attribute level
V(di)$level

# Extract vertex attribute levelName_full
V(di)$levelName_full

# Extract vertex attribute levelName
V(di)$levelName

# Extract edge attribute weight
E(di)$weight

# Extract edge attribute layer
E(di)$layer
```
</details>

Next in [Vignette 3 - Weight](Vignette-3---Weight.html), you will find out about weighting networks.

## Last updated
This vignette is associated with `AHgen` v1.0.0, and was last updated by Melissa Bedinger on `r Sys.Date()`.