---
title: "Application - <br>Compare scenarios"
author: 
 - Annie Visser-Quinn & Melissa Bedinger
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: paper
    highlight: haddock
vignette: >
  %\VignetteIndexEntry{Vignette 7E - Application - Compare scenarios}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)

rm(list = ls()); cat("\014")

library(knitr)
library(tidyverse)
library(AHgen)
```

<style>
#TOC {
  background: url("images/hex.png");
  background-size: 220px;
  background-position-x: center;
  background-position-y: top;
  padding-top: 280px !important;
  background-repeat: no-repeat;
}
</style>

---

In this vignette, you will go through the steps of comparing Urban Systems Abstraction Hierarchies in `AHgen`:

1. Prepare the environment
2. Compare scenarios to a single benchmark
3. Compare scenarios to respective (multiple) benchmarks
4. Export

To navigate to other vignettes, go to [Vignette 0 - Welcome to AHgen](Vignette-0---Welcome-to-AHgen.html).


## 1. Prepare the environment
```{r, eval = FALSE}
# Clear the RStudio environment
rm(list = ls()); cat("/014"); gc(verbose = TRUE)

# CTRL + SHIFT + F10 will detach any loaded packages and restart R

# Set the working directory to the folder where this script is saved
setwd(dirname(rstudioapi::getSourceEditorContext()$path)); getwd()

# If you are doing this for the first time

# install the package pacman which checks to see if a package is installed
# and if not it attempts to install the package from CRAN 
# and/or any other repository in the pacman repository list

# and install the package devtools which will allow you to download a package straight from GitHub 
# i.e. if they are not yet on CRAN and/or the pacman repository list

install.packages("devtools")
install.packages("pacman")

# Ensure latest version of AHgen is installed
devtools::install_github("avisserquinn/AHgen", dependencies = TRUE)

# If package has 'non-zero exit status', try re-installing devtools & rlang
install.packages("devtools")
install.packages("rlang")

# Load required packages
pacman::p_load(tidyverse, installr, janitor, AHgen)
```


## 2. Compare scenarios to a single benchmark

If you want to compare a single new scenario to a single benchmark, you can adapt the code below.
```{r}
# Read in the AH_benchmark you want to use as a point of comparison
USAH_benchmark =
  read_rds("inputs/toCompare/USAH_3.0_Bristol_baseline_20230719-110748.RDS")

# Create a designated folder (e.g. "toCompare") 
# where you've copied all individual scenario outputs you are comparing to the benchmark

# Read in all the file names from the designated folder
scenarioFilenames = 
  list.files(
    path = "inputs/toCompare", 
    pattern = ".*USAH_.*\\.RDS", 
    full.names = TRUE)

# Read in all scenario files to a list of lists
scenarios_toCompare = 
  lapply(scenarioFilenames, function (x) readRDS(x))

# Create vector for file names without path and extension
scenarioNames = 
  list.files(
    path = "inputs/toCompare", 
    pattern = ".*USAH_.*\\.RDS") %>% 
  tools::file_path_sans_ext()

# Set names of allScenarios list elements to reflect correct scenario
names(scenarios_toCompare) = scenarioNames

# Compare the Abstraction Hierarchies
Bristol_floodRiver_compared = 
  compare_AH(type = "USAH",
             AH_benchmark = USAH_benchmark, 
             scenarios_toCompare = scenarios_toCompare, 
             scenarioNames = scenarioNames)

# Inspect the compared vertices
Bristol_floodRiver_compared$vertices

# Inspect the compared edges
Bristol_floodRiver_compared$edges

# Inspect the compared vExcluded
Bristol_floodRiver_compared$vExcluded

# Inspect the compared results
Bristol_floodRiver_compared$results

# Inspect the compared confidence ratings
# if a sensitivity analysis has been applied to any of the input scenarios
Bristol_floodRiver_compared$confidence
```


## 3. Compare scenarios to respective (multiple) benchmarks

If you have compiled several comparisons similar to Section 2, you can also combine results to compare multiple scenarios to their respective (multiple) benchmarks, as below.

```{r}
# Read in the data
# individual comparisons of a flood scenario to a benchmark scenario for five different locations
Bristol =   
  read_rds(
    "inputs/comparison_USAH_3.0_Bristol-baseline_Bristol-floodRiver-100_20230720-173152.RDS")

Edinburgh = 
  read_rds(
    "inputs/comparison_USAH_3.0_Edinburgh-baseline_Edinburgh-floodRiver-200_20230720-173512.RDS")

Glasgow = 
  read_rds(
    "inputs/comparison_USAH_3.0_Glasgow-baseline_Glasgow-floodRiver-200_20230720-173627.RDS")

LondonCentral = 
  read_rds(
    "inputs/comparison_USAH_3.0_London-Central-baseline_London-Central-floodRiver-100_20230720-173814.RDS")

Manchester = 
  read_rds(
    "inputs/comparison_USAH_3.0_Manchester-baseline_Manchester-floodRiver-100_20230720-173929.RDS")


# Combine the five comparisons
fiveCities_floodRiver_combined = 
  Bristol %>%
  purrr::map2(Edinburgh, rbind) %>%
  purrr::map2(Glasgow, rbind) %>%
  purrr::map2(LondonCentral, rbind) %>%
  purrr::map2(Manchester, rbind)

# Generate a new summary of confidence ratings
fiveCities_floodRiver_combined$confidence =
  fiveCities_floodRiver_combined$results %>%
  summarise_confidence()

# Inspect the compared vertices
fiveCities_floodRiver_combined$vertices

# Inspect the compared edges
fiveCities_floodRiver_combined$edges

# Inspect the compared vExcluded
fiveCities_floodRiver_combined$vExcluded

# Inspect the compared results
fiveCities_floodRiver_combined$results

# Inspect the compared confidence ratings
fiveCities_floodRiver_combined$confidence
```


## Last updated
This vignette is associated with `AHgen` v1.0.0, and was last updated by Melissa Bedinger on `r Sys.Date()`.