---
title: "Welcome to AHgen"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: paper
    highlight: haddock
vignette: >
  %\VignetteIndexEntry{Vignette 1 - Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(
  root.dir = normalizePath('../../AHgen')
  )

rm(list = ls()); cat("\014")

library(knitr)
library(tidyverse)
library(igraph)
```

AHgen is **under development**, V0.0.1 is available as a pre-release here: https://github.com/avisserquinn/AHgen/releases/tag/V0.0.1

## What is an abstraction hierarchy?
The Abstraction Hierarchy enables analysis of systems at any scale – from microwaves to transport systems to entire cities. By understanding how physical objects connect to different – yet interdependent – functions, we can improve the design of the system.

<span style = "color:red;">Let’s consider some key terms before we continue.</span>

- A **network** is a collection of parts (**vertices**) and connections between parts (**edges**)
- The **Abstraction Hierarchy** is a type of **hierarchical network** that **connects** the **tangible** to the **intangible** 

In the Abstraction Hierarchy:

- Vertices are structured across five **levels** of increasing abstraction, from physical objects to the functional purpose(s) of the system
- Edges, the links between two levels (a **layer**), representing the functional connections between the physical and abstract

<span style = "color:red;">Abstraction hierarchy image(s)?</span>

<br>

## What is AHgen?
Applications of the Abstraction Hierarchy typically aim to visualise the network, relying on pen-and-paper methods, PowerPoint drawing, or proprietary software. AHgen brings the Abstraction Hierarchy into the 21st century by providing an open source software code in R. 

<span style = "color:red;">Other advantages to AHgen include:</span>

- User-friendly and accessible, it does not require previous R or coding experience
- Designed using a Building Block Method, AHgen may be readily extended by the more advanced R user 

<br>

## AHgen workflow and functions

AHgen is extremely flexible, with lots of possible workflows. There are five families of functions: 

- **Input** Reading in the Abstraction Hierarchy and accompanying data
- **Convert** Converting the Abstraction Hierarchy into a various formats compatible with network analysis
- **Weight** Weighting the edges
- **Analyse** Applying network analysis
- **Visualise** An innovative way of visualising the Abstraction Hierarchy as a circle

<span style = "color:red;">Melissa figure here</span>

There are four vignettes designed to get you on your way: <span style = "color:red;">Add links</span>

- Vignette 1 
- Vignette 2
- Vignette 3
- Vignette 4

<br>

## Applications
Three walkthrough vignettes are provided to illustrate the application of AHgen across various domains and scales:

<span style = "color:red;">Make these into links but also maybe have a line on what type of domain/scale they cover</span>

- Vignette 5
- Vignette 6
- Vignette 7

<span style = "color:red;">Not sure what to do with this text right now.</span>

AHgen can be applied to Abstraction Hierarchies in any domain at any scale for various objectives. The Urban Systems Abstraction Hierarchy (USAH) is an example of an Abstraction Hierarchy on a massive scale. Exploring USAH with AHgen it’s possible to:

- Look at the interdependencies between urban functions
- Explore how these functions are impacted by different conditions
- Explore the impact of hazards by degrading connections in the system (edge weighting)
- Explore the impact of reconnecting vertices and edges (effectively redesigning the system)

<br>

## Getting started {.tabset}
<span style = "color:red;">This might be better off in its own section.</span>

To get started with AHgen, you'll need to: 

- Step 1 - **Install** R and (optional) RStudio
- Step 2 - Install the required R **packages**
- Step 3 - **Download the AHgen** source code from this GitHub repository
- Step 4 - Prepare a script in R to **load the AHgen functions**

A walkthrough of each step is provided via the **tabs** below.


### Step 1 - Install

These screenshots will guide you through the process of downloading [R](https://cloud.r-project.org/) and [RStudio](https://rstudio.com/products/rstudio/download/). But first:

- **What is R?** - R is an open source (free) programming language used to develop statistical software such as AHgen. 
- **What is RStudio?** - RStudio is an "integrated development environment for R". In simpler terms, RStudio provides a more user-friendly environment to run R code from. It is also open source and highly recommended. 

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.align = "center"}
files <- list.files("vignettes/images/gettingStarted", pattern = ".png", full.names = TRUE)

include_graphics(normalizePath(files))
```

### Step 2 - R packages

**Required packages:** `readxl`, `openxlsx`, `tidyverse`, `igraph`, `tnet` and `ggraph`

An **R package** is a collection of functions and/or data that have been *packaged* together. For more information, check out [this handy guide](https://www.datacamp.com/community/tutorials/r-packages-guide). 

To install the required packages:

- Open RStudio
- Paste `install.packages(c("readxl", "openxlsx", "tidyverse", "igraph", "tnet", "ggraph"))` into the console (bottom left hand side) and press `enter`. *Note, you might be asked to install dependencies or update packages*

A package is like a book, you use a `library()` to check a package out of the library. AHgen will load these packages for you, or you can copy and paste the code chunk below.

```{r, eval = FALSE}
library(readxl)
library(openxlsx)
library(tidyverse)
library(igraph)
library(tnet)
library(ggraph)
```

### Step 3 - Download AHgen
<span style="color: red;">To be filled in once the github is more up to date.</span>

### Step 4 - Load AHgen functions
The functions that make up AHgen are located in the *functions* folder. You can load all of them in by running the script in *functions.R* using `source()`. Step-by-step: 

- Open RStudio
- Navigate to *File > New file > R Script*
- Save the script in the AHgen directory (folder)
- Copy and paste the code chunk below into your script

```{r, eval = FALSE}
rm(list = ls()) # Clear environment
cat("\014") # Clear console

setwd(dirname(rstudioapi::getSourceEditorContext()$path)) # Set working directory to script directory

source("functions/functions.R") # Read in the functions script
```

- Highlight your script and run by pressing `Ctrl+Enter`. The names of the functions will appear in the top right corner
- Type `ls()` directly in the console to print the names of the loaded functions, shown below

```{r, include = FALSE}
source("functions/functions.R")
```
```{r, echo = FALSE}
ls()
```

### {-}

<br>

## Acknowledgements
The AHgen was developed to compare Urban Systems Abstraction Hierarchy models (USAH) for UK cities. Development is funded by the UKRI EPSRC as part of the *Water Resilient Cities project*, grant number EP/N030419/1, as well as an associated EPSRC Impact Acceleration Fund, *Urban system response and recovery: Development of an online tool to explore the impact of COVID-19*.


## Contributions & citation
**Contributions**

<span style="color: red;">*Conceptualisation & development*: Melissa Bedinger *Lead developer:* Annie Visser-Quinn *What other ones do we want to put here?* </span>

**Contact** If you have any questions about using or running the code, please contact m.bedinger@hw.ac.uk. 

**Please cite** Annie Visser-Quinn and Melissa Bedinger. (2020) AHgen V0.0.1, Edinburgh, Scotland. doi:10.5281/zenodo.3975549 [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3975549.svg)](https://doi.org/10.5281/zenodo.3975549) <span style="color: red;">To be updated once we can do a new Zenodo</span>