---
title: "Getting started"
author:  
 - Annie Visser-Quinn & Melissa Bedinger
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: paper
    highlight: haddock
vignette: >
  %\VignetteIndexEntry{1 - Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE)

rm(list = ls()); cat("\014")

library(knitr)
library(tidyverse)
library(AHgen)
```

<style>
#TOC {
  background: url("images/hex.png");
  background-size: 220px;
  background-position-x: center;
  background-position-y: top;
  padding-top: 280px !important;
  background-repeat: no-repeat;
}
</style>

---

In this vignette, you will install everything you need to get started with `AHgen`:

- **Step 1** - Install **R** and **RStudio**
- **Step 2** - Install the required **R packages**
- **Step 3** - Install **`AHgen`** from a GitHub repository

Each step is detailed via the tabs below.

To navigate to other vignettes, go to [Vignette 0 - Welcome to AHgen](Vignette-0---Welcome-to-AHgen.html).

## Install guide {.tabset}

### **Step 1 - Install R & RStudio**

#### What is R?
R is an open source (free) programming language used to develop statistical software. 

#### What is RStudio?
RStudio is an "integrated development environment for R". In simpler terms, RStudio provides a more user-friendly environment to run R code from. It is also open source and highly recommended.

#### How do I install R & RStudio?

The screenshots below will guide you through the process of downloading [R](https://cloud.r-project.org/) and [RStudio](https://rstudio.com/products/rstudio/download/). Note that the versions of R and RStudio shown in the screenshots below will now be outdated; please download the most recent version.
<br>

```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/1 - Getting started/Step 1/1.png"),
               alt = 'install', width = 700)
```
<br>

```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/1 - Getting started/Step 1/2.png"),
               alt = 'install-complete', width = 700)
```
<br>

```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/1 - Getting started/Step 1/3.png"),
               alt = 'install', width = 700)
```
<br>

```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/1 - Getting started/Step 1/4.png"),
               alt = 'install-complete', width = 700)
```
<br>

```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/1 - Getting started/Step 1/5.png"),
               alt = 'install-complete', width = 700)
```
<br>

```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/1 - Getting started/Step 1/6.png"),
               alt = 'install-complete', width = 700)
```
<br>

```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/1 - Getting started/Step 1/7.png"),
               alt = 'install-complete', width = 700)
```
<br>

### **Step 2 - Install R packages**

#### What is an R package?
An R package is a collection of functions and/or data that have been packaged together. For more information, check out [this guide](https://www.datacamp.com/community/tutorials/r-packages-guide).

#### Packages to install other packages
There are many ways to install R packages. In example `AHgen` scripts, there are two packages which are used for the specific purpose of installing other packages. We find these two packages make the application of `AHgen` smoother.

- [**`devtools`**](https://cran.r-project.org/web/packages/devtools/readme/README.html) - This package makes package development easier by providing R functions that simplify and expedite common tasks. The [`install_github()`](https://www.rdocumentation.org/packages/devtools/versions/1.13.6/topics/install_github) function is used in example `AHgen` scripts, to install a package straight from GitHub. This is useful if a package is not yet hosted on CRAN (such as `AHgen`), or if you want to play with a different [branch](https://r-bio.github.io/intro-git-rstudio/) of a package (e.g. the newest 'development' version).
- [**`pacman`**](https://cran.r-project.org/web/packages/pacman/readme/README.html) - This package is an R package management tool that combines the functionality of base library related functions into intuitively named functions. The [`p_load()`](https://www.rdocumentation.org/packages/pacman/versions/0.5.1/topics/p_load) function is used in example `AHgen` scripts, to load multiple packages at the same time instead of using the library() function. Any packages in the `p_load()` function will also be installed if you have not already done so; this acts as an extra check.


#### AHgen dependencies

`AHgen` uses some functions from [**`base` R**](https://cran.r-project.org/doc/FAQ/R-FAQ.html#R-Add_002dOn-Packages) which do not require you to install any additional packages. In particular, `AHgen` uses [`base`](https://stat.ethz.ch/R-manual/R-devel/library/base/html/00Index.html), [`grid`](https://stat.ethz.ch/R-manual/R-devel/library/grid/html/00Index.html), [`grDevices`](https://stat.ethz.ch/R-manual/R-devel/library/grDevices/html/00Index.html), and [`stats`](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/00Index.html). 

`AHgen` also relies on functions in several other pre-existing packages. These are called **dependencies**.

##### *Dependencies*

- the [**`tidyverse`**](https://www.tidyverse.org/) is actually a *collection* of R packages designed for data science. All packages share an underlying design philosophy, grammar, and data structures.


- [**`Biobase`**](https://bioconductor.org/packages/release/bioc/html/Biobase.html) and [`BiocManager`](https://www.bioconductor.org/install/) implements base functions for `Bioconductor`. It has a helpful function called [`subListExtract()`](https://www.rdocumentation.org/packages/Biobase/versions/2.32.0/topics/subListExtract) to extract the same element from the sublists of a list. For example, if outputs of `AHgen` are in [list format](https://www.geeksforgeeks.org/r-lists/), and every `AHgen` output has a list element called 'vertices', `subListExtract()` can extract the list element 'vertices' from multiple lists quickly. This is useful for comparing `AHgen` outputs. To install `Biobase` you will first need to install the package `BiocManager`. This package serves as the primary way to ensure that the appropriate `Biobase` installation is used with respect to the version of R in use, regardless of the R and `Bioconductor` release cycles.
- [**`data.table`**](https://cran.r-project.org/web/packages/data.table/readme/README.html) provides a high-performance version of base R’s `data.frame`, with syntax and feature enhancements for ease of use, convenience and programming speed.
- [**`igraph`**](https://cran.r-project.org/web/packages/igraph/index.html) provides routines for simple graphs and network analysis.
- [**`installr`**](https://cran.r-project.org/web/packages/installr/readme/README.html) offers a set of R functions for the installation and updating of software.
- [**`janitor`**](https://cran.r-project.org/web/packages/janitor/readme/README.html) provides simple functions for examining and cleaning dirty data.
- [**`vctrs`**](https://cran.r-project.org/web/packages/vctrs/readme/README.html) provides tools to ensure that functions behave consistently with respect to inputs of varying length and type.
- [**`xlsx`**](https://cran.r-project.org/web/packages/xlsx/readme/README.html) provides R functions to read, write, and format Excel files.

- [**`ggh4x`**](https://cran.r-project.org/web/packages/ggh4x/readme/README.html),  [**`gginnards`**](https://cran.r-project.org/web/packages/gginnards/readme/README.html), [**`ggraph`**](https://cran.r-project.org/web/packages/ggraph/readme/README.html), [**`ggrepel`**](https://cran.r-project.org/web/packages/ggrepel/readme/README.html), and [**`scales`**](https://cran.r-project.org/web/packages/scales/readme/README.html) are all [`ggplot2`](https://ggplot2.tidyverse.org/) extension packages to customise visualisations.
- [**`plotly`**](https://cran.r-project.org/web/packages/plotly/readme/README.html) helps to create interactive web graphics via the open source JavaScript graphing library [plotly.js](https://github.com/plotly/plotly.js).
- [**`see`**](https://cran.r-project.org/web/packages/see/) provides plotting utilities supporting packages in the [`easystats`](https://cran.r-project.org/web/packages/easystats/index.html) ecosystem  and some extra themes, geoms, and scales for [`ggplot2`](https://ggplot2.tidyverse.org/)
- [**`sf`**](https://cran.r-project.org/web/packages/sf/index.html) provides support for simple features, a standardized way to encode spatial vector data.
- [**`tidytext`**](https://cran.r-project.org/web/packages/tidytext/readme/README.html) provides functions and supporting data sets to allow conversion of text to and from tidy formats, and to switch seamlessly between tidy tools and existing text mining packages.

Some packages are needed to borrow only a single function; other packages are heavily relied upon. For more detail on dependencies you can see *AHgen/data-raw/functionsTracker_20230707.xlsx*.


##### Install R packages

To install `devtools` and `pacman`:

1. Open RStudio. 
2. Paste `install.packages(c("devtools", "pacman"))` into the console (bottom left hand side).
3. Press `enter`. 

It should look something like below.

```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/1 - Getting started/Step 2/1.png"),
               alt = 'install', width = 700)
```
<br>
```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri("images/1 - Getting started/Step 2/2.png"),
               alt = 'install-complete', width = 700)
```
<br>

To install `BiocManager` and `Biobase`, use the following code.

```{r, eval = FALSE}
# Install BiocManager and Biobase 
# Following instructions from https://www.bioconductor.org/packages/release/bioc/html/Biobase.html

if (!requireNamespace("BiocManager", quietly = TRUE))
install.packages("BiocManager")
BiocManager::install("Biobase")
```


To install the rest of the dependencies, use the following code.

```{r, eval = FALSE}
# Install AHgen dependencies
pacman::p_load(
  tidyverse, data.table, igraph, installr, janitor, vctrs, xlsx,
  ggh4x, gginnards, ggraph, ggrepel, scales, plotly, see, sf, tidytext)
```
<br>

### **Step 3 - Install `AHgen`**

To install the `AHgen` package itself, use the following code.

The argument `dependencies = TRUE` will include installation of any `AHgen` dependencies i.e. supporting packages - in case they have not already been installed.

```{r, eval = FALSE}
devtools::install_github("avisserquinn/AHgen", dependencies = TRUE)
```

And you're ready to go!

If you want to explore, load the `AHgen` package.
```{r}
pacman::p_load(AHgen)
```

See a full list of its current functions.
```{r}
# List the names of AHgen functions and data
ls("package:AHgen")

# List the AHgen function names and required inputs
lsf.str("package:AHgen")
```
<br>


## Next
Next in [Vignette 2 - Input & Convert](Vignette-2---Input-&-Convert.html), you will find out about inputs and data in `AHgen`.

## Last updated
This vignette is associated with `AHgen` v1.0.0, and was last updated by Melissa Bedinger on `r Sys.Date()`.