---
title: "Vignette 3 - Analysis"
author: 
 - Dr Annie Visser-Quinn (a.visser-quinn@hw.ac.uk)
 - Dr Melissa Bedinger (m.bedinger@hw.ac.uk)
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: paper
    highlight: haddock
vignette: >
  %\VignetteIndexEntry{Vignette 3 - Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE, 
                      message=FALSE, 
                      warning=FALSE)

knitr::opts_knit$set(
  root.dir = normalizePath('../../AHgen')
  )

rm(list = ls()); cat("\014")

library(knitr)
library(tidyverse)
library(igraph)
```

```{r, echo = FALSE}
htmltools::img(src = knitr::image_uri(file.path("vignettes/images/hex.png")), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px; max-width:20%')
```

```{r include = FALSE}
source("functions/functions.R")

dh <- read_adjMat("inputs/AH_mobility.xlsx") 
dv <- dh %>% select(level, levelName, vName)
de <- dh %>% adjMat_to_edgelist(dv)
di <- dh %>% adjMat_to_igraph(dv)
di_w <- weightEdges(de, read_xlsx("inputs/exampleWeights.xlsx")) %>% edgelist_to_igraph(dv)


# Workaround for Rmarkdown filepath errors later in script
source("functions/functions_internal_calc.R") 

calcWVBC2 <- function(igraph, vInfo) {

  require(tnet)
  
  tnet <- 
    igraph %>% 
    as_adjacency_matrix(attr = "weight") %>%
    as.matrix
  
  tnet2 <- 
    tnet %>%   
    symmetrise_w() %>%
    betweenness_w(directed = TRUE, alpha = 0.5) %>%
    as_tibble %>%
    mutate(vName = rownames(tnet)) %>%
    select(vName, WVBC = betweenness)
  
  output <- function_merge_vInfo(dt = tnet2, vInfo = vInfo)

  return(output)
  
}

calcUWVBC2 <- function(igraph, vInfo) { 

  require(igraph)
  
  igraph %>% 
    remove.edge.attribute("weight") %>%
    igraph::betweenness() %>%
    function_igraphResultFormatting(name = "UWVBC") %>%
    function_merge_vInfo(vInfo)
  
}
```

<span style="color: red;">@Melissa Explain what type of analysis is done for abstraction hierarchy. Explain centralities (as a minimum).</span>

## 1. (Vertex) Betweenness centrality

AHgen includes two centrality functions:

- Unweighted vertex betweenness centrality (UWVBC) calculated using `calcUWVBC()`
- Weighted vertex betweenness centrality (WVBC) calculated using `calcWVBC()`

Both require two input arguments, an igraph and the vertex information.

Click below for example applications.

<details><summary><span style = "color: #2196F3;">**Example 1 - UWVBC**</span></summary>


The calculation of UWVBC effectively ignores the weight attribute. 

For this example, let's revisit the mobility abstraction hierarchy - we'll look at one igraph where the weights are all equal to 1, and one where we considered an aging vehicle.

First let's look at the case where all the weights are the same.

```{r}
E(di)$weight
di %>% calcUWVBC(dv)
```

And now the case where we weight the edges linked to the vehicle component vertex.

```{r}
E(di_w)$weight
di_w %>% calcUWVBC(dv)
```

As you can see, the UWVBC for both is the same.

</details>

<details><summary><span style = "color: #2196F3;">**Example 2 - WVBC**</span></summary>

Let's continue with the same input data and explore the `calcWVBC()` function. We'll look at the same two igraphs again. Now that the centrality is based on weights, we expect to see different values for each. 
```{r}
di %>% calcWVBC(dv)
di_w %>% calcWVBC(dv)
```

</details>

## 2. Calculating multiple metrics

What if you want to compare the UWVBC and WVBC functions side by side? You can use AHgen's helper function `calcMetrics()`. 

`calcMetrics()` has one additional input argument: a vector of the names of the metrics to be calculated, e.g. `c("UWVBC", "WVBC").

Currently only the unweighted and weighted vertex betweenness centralities can be calculated this way. To expand `calcMetrics()`, check out section 3.

<details><summary><span style = "color: #2196F3;">**Example 3**</span></summary>

Let's use `calcMetrics()` to repeat examples 1 and 2 in a tidier way.

```{r, eval = FALSE}
calcMetrics(di, dv, c("UWVBC", "WVBC"))
```
```{r, echo = FALSE}
output <- list(calcMetrics(di, dv, "UWVBC"),
               calcMetrics(di, dv, "WVBC"))

output <- 
  output %>% 
  reduce(full_join, by = c("level", "levelName", "vName")) %>%
  arrange(level, levelName, vName)

output
```

```{r, eval = FALSE}
calcMetrics(di_w, dv, c("UWVBC", "WVBC"))
```
```{r echo = FALSE}
output <- list(calcMetrics(di_w, dv, "UWVBC"),
               calcMetrics(di_w, dv, "WVBC"))

output <- 
  output %>% 
  reduce(full_join, by = c("level", "levelName", "vName")) %>%
  arrange(level, levelName, vName)

output
```

</details>

## 3. Adding additional metrics
It's possible to expand `calcMetrics()` to compute additional metrics by defining functions which follow the same structure as `calcUWVBC()`. To check out the structure of a function simply type it's name without the brackets

```{r}
calcUWVBC
```
The outputs from the igraph functions `betweenness()` and `eigen_centrality()` are the same format. Therefore we can create a function to calculate Eigen centrality by changing the code in three places:

- Changing the name. Make sure the name is prefixed with *calc* otherwise `calcMetrics()` will not be able to call it
- Replacing `igraph %>% igraph::betweenness()` with `igraph::eigen_centrality(igraph)$vector`
- Changing the internal column name from *UWVBC* to *Eigen*

Check out the example below to see it in action.

<details><summary><span style = "color: #2196F3;">**Example 4**</span></summary>

```{r}
calcEigen <- function(igraph, vInfo) { 
  
  source("functions/functions_internal_calc.R")
  
  require(igraph)
  
  igraph::eigen_centrality(igraph)$vector %>%
    function_igraphResultFormatting(name = "Eigen") %>%
    function_merge_vInfo(vInfo)
  
}

calcEigen(di_w, dv)
```
Now that we've defined the function we can use it within `calcMetrics()`.

```{r, eval = FALSE}
calcMetrics(di_w, dv, c("UWVBC", "WVBC", "Eigen"))
```
```{r echo = FALSE}
output <- list(calcMetrics(di_w, dv, "UWVBC"),
               calcMetrics(di_w, dv, "WVBC"),
               calcMetrics(di_w, dv, "Eigen"))

output <- 
  output %>% 
  reduce(full_join, by = c("level", "levelName", "vName")) %>%
  arrange(level, levelName, vName)

output
```

</details>

## 4. Calculate change signal
To calculate the percent change from one igraph to another, for any given metric, we can use the AHgen function `calcChange()`. There are three input arguments:

- `before` A metric from the initial igraph
- `after` A metric from the modified/weighted/impacted igraph
- `metric` The column name of the metric for which change is to be calculated

<details><summary><span style = "color: #2196F3;">**Example 5**</span></summary>

Let's revisit the same igraphs from before, `di` and `di_w`. First, we need to calculate the centralities.

```{r}
before <- di %>% calcWVBC(dv)
after <- di_w %>% calcWVBC(dv)

before
after
```
 
Now we can calculate the change signal using `calcChange()`.

```{r}
calcChange(before, after, "WVBC")
```

You could also embed the calculation of the before and after centralities inside the function.

```{r}
calcChange(before = di %>% calcWVBC(dv),
           after = di_w %>% calcWVBC(dv),
           metric = "WVBC")
```