---
title: "Application - <br>Introduce other hazards"
author: 
 - Annie Visser-Quinn & Melissa Bedinger
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: paper
    highlight: haddock
vignette: >
  %\VignetteIndexEntry{Vignette 7D - Application - Introduce other hazards}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)

rm(list = ls()); cat("\014")

library(knitr)
library(tidyverse)
library(AHgen)
```

<style>
#TOC {
  background: url("images/hex.png");
  background-size: 220px;
  background-position-x: center;
  background-position-y: top;
  padding-top: 280px !important;
  background-repeat: no-repeat;
}
</style>

---

In this vignette, you will go through the steps of applying other hazards to an Urban Systems Abstraction Hierarchy template in `AHgen`:

1. Prepare the environment
2. Read in the data
3. Generate and apply a hypothetical drought hazard
4. Generate and apply a hypothetical flood hazard
5. Generate and apply a hypothetical pandemic hazard
6. Export

To navigate to other vignettes, go to [Vignette 0 - Welcome to AHgen](Vignette-0---Welcome-to-AHgen.html).


## 1. Prepare the environment

```{r, eval = FALSE}
# Clear the RStudio environment
rm(list = ls()); cat("/014"); gc(verbose = TRUE)

# CTRL + SHIFT + F10 will detach any loaded packages and restart R

# Set the working directory to the folder where this script is saved
setwd(dirname(rstudioapi::getSourceEditorContext()$path)); getwd()

# If you are doing this for the first time

# install the package pacman which checks to see if a package is installed
# and if not it attempts to install the package from CRAN 
# and/or any other repository in the pacman repository list

# and install the package devtools which will allow you to download a package straight from GitHub 
# i.e. if they are not yet on CRAN and/or the pacman repository list

install.packages("devtools")
install.packages("pacman")

# Ensure latest version of AHgen is installed
devtools::install_github("avisserquinn/AHgen", dependencies = TRUE)

# If package has 'non-zero exit status', try re-installing devtools & rlang
install.packages("devtools")
install.packages("rlang")

# Load required packages
pacman::p_load(tidyverse, installr, janitor, AHgen)
```


## 2. Read in the data

```{r}
# Set names for AH, version, location, and scenarios
AH_name = "USAH"
AH_version = "3.0"
AH_location = "template"
scenario_drought = "drought-hypothetical"
scenario_flood = "flood-hypothetical"
scenario_pandemic = "pandemic-hypothetical"

# Set directory for outputs
directory = "outputs/"

# Read in the USAH input you want to start with
USAH_template_baseline = 
  read_rds("inputs/USAH_3.0_template_baseline_20230719-104712.RDS")

# Extract input edgelist
edgelist_template_baseline = USAH_template_baseline$edgelist

# Read in a guide specifying which vertices may be affected by a hypothetical drought
vDrought = 
  readxl::read_xlsx("inputs/vGuide_2.0_template_drought-hypothetical_20210720.xlsx")

# Read in a guide specifying which edges may be affected by a hypothetical flood
eFlood = 
  readxl::read_xlsx("inputs/edgeGuide_2.0_template_flood-hypothetical_20210803.xlsx")

# Read in a guide specifying which edges may be affected by a hypothetical pandemic
ePandemic = 
  readxl::read_xlsx("inputs/edgeGuide_2.0_template_pandemic-hypothetical_20210805.xlsx")

# Inspect the data
USAH_template_baseline
vDrought
eFlood
ePandemic
```


## 3. Generate and apply a hypothetical drought hazard
```{r}
# Create vector of Process vertices affected by a drought
vDrought = vDrought %>% pull(Node)

# Create edgesDrought
edgesDrought = 
  edgelist_template_baseline %>% 
  filter(from %in% vDrought) %>%
  mutate(weight = 0.5) %>%
  rename(weightNew = weight)

# Inspect new edges representing the drought hazard to be introduced
edgesDrought

# Check for duplicate edges (that may have been accidentally added to a 
# manual edgelist for a hypothetical flood scenario) - confirms whether there 
# are any potential conflicts to sort out before introducing the new edge 
# weights for that scenario
edgesDrought %>% check_duplicates()

# If any, inspect the duplicate edges with (potentially) different weights
# and correct input file then re-import before proceeding

# Checking for outdated or inconsistent edges in the edge guide 
# that are inconsistent with the USAH_input edgelist you want to overwrite 
# both in terms of outdated linkages and outdated vertex names,
# this prevents adding anything odd to the structure of the network
edgesDrought %>% check_diff(edgelist = edgelist_template_baseline)

# If any, inspect the edgesNew that do not match and correct input file 
# then re-import before proceeding

# Introduce edgesDrought to create edgelist_template_drought
edgelist_template_drought = 
  edgelist_template_baseline %>%
  weight_edges(edgesDrought, remove = FALSE) %>%
  weight_hangingVertices(remove = FALSE)

# Generate output for USAH_template_drought
USAH_template_drought = 
  apply_scenario(
    AH_input = USAH_template_baseline, 
    edgelist_scenario = edgelist_template_drought,
    AH_name = AH_name, 
    AH_version = AH_version, 
    AH_location = AH_location, 
    AH_scenario = scenario_drought)

# Inspect removed nodes
USAH_template_drought$vExcluded

# Inspect results
USAH_template_drought$results
```


## 4. Generate and apply a hypothetical flood hazard
```{r}
# Select only necessary columns
edgesFlood = eFlood %>% select(layer, from, to, weightNew)
  
# Inspect new edges representing the flood hazard to be introduced
edgesFlood

# Check for duplicate edges (that may have been accidentally added to a 
# manual edgelist for a hypothetical flood scenario) - confirms whether there 
# are any potential conflicts to sort out before introducing the new edge 
# weights for that scenario
edgesFlood %>% check_duplicates()

# If any, inspect the duplicate edges with (potentially) different weights
# and correct input file then re-import before proceeding

# Checking for outdated or inconsistent edges in the edge guide 
# that are inconsistent with the USAH_input edgelist you want to overwrite 
# both in terms of outdated linkages and outdated vertex names,
# this prevents adding anything odd to the structure of the network
edgesFlood %>% check_diff(edgelist = edgelist_template_baseline)

# If any, inspect the edgesNew that do not match and correct input file 
# then re-import before proceeding

# Introduce edgesFlood to create edgelist_template_flood
edgelist_template_flood = 
  edgelist_template_baseline %>%
  weight_edges(edgesFlood, remove = FALSE) %>%
  weight_hangingVertices(remove = FALSE)

# Generate output for USAH_template_flood
USAH_template_flood = 
  apply_scenario(
    AH_input = USAH_template_baseline, 
    edgelist_scenario = edgelist_template_flood,
    AH_name = AH_name, 
    AH_version = AH_version, 
    AH_location = AH_location, 
    AH_scenario = scenario_flood)

# Inspect removed nodes
USAH_template_flood$vExcluded

# Inspect results
USAH_template_flood$results
```


## 5. Generate and apply a hypothetical pandemic hazard
```{r}
# Select only necessary columns
edgesPandemic = ePandemic %>% select(layer, from, to, weightNew)
  
# Inspect new edges representing the pandemic hazard to be introduced
edgesPandemic

# Check for duplicate edges (that may have been accidentally added to a 
# manual edgelist for a hypothetical pandemic scenario) - confirms whether there 
# are any potential conflicts to sort out before introducing the new edge 
# weights for that scenario
edgesPandemic %>% check_duplicates()

# If any, inspect the duplicate edges with (potentially) different weights
# and correct input file then re-import before proceeding

# Checking for outdated or inconsistent edges in the edge guide 
# that are inconsistent with the USAH_input edgelist you want to overwrite 
# both in terms of outdated linkages and outdated vertex names,
# this prevents adding anything odd to the structure of the network
edgesPandemic %>% check_diff(edgelist = edgelist_template_baseline)

# If any, inspect the edgesNew that do not match and correct input file 
# then re-import before proceeding

# Introduce edgesPandemic to create edgelist_template_pandemic
edgelist_template_pandemic = 
  edgelist_template_baseline %>%
  weight_edges(edgesPandemic, remove = FALSE) %>%
  weight_hangingVertices(remove = FALSE)

# Generate output for USAH_template_pandemic
USAH_template_pandemic = 
  apply_scenario(
    AH_input = USAH_template_baseline, 
    edgelist_scenario = edgelist_template_pandemic,
    AH_name = AH_name, 
    AH_version = AH_version, 
    AH_location = AH_location, 
    AH_scenario = scenario_pandemic)

# Inspect removed nodes
USAH_template_pandemic$vExcluded

# Inspect results
USAH_template_pandemic$results
```


## 6. Export
```{r, eval = FALSE}
USAH_template_drought %>% 
  export_AHgen(
    type = "USAH",
    directory = directory,
    AH_version = AH_version, 
    AH_location = AH_location, 
    AH_scenario = scenario_drought)

USAH_template_flood %>% 
  export_AHgen(
    type = "USAH",
    directory = directory,
    AH_version = AH_version, 
    AH_location = AH_location, 
    AH_scenario = scenario_flood)

USAH_template_pandemic %>% 
  export_AHgen(
    type = "USAH",
    directory = directory,
    AH_version = AH_version, 
    AH_location = AH_location, 
    AH_scenario = scenario_pandemic)
```


## Next
Next in [Vignette 7E - Application - Compare scenarios](Vignette-7E---Application---Compare-scenarios.html), you will go through the steps of comparing Urban Systems Abstraction Hierarchies in `AHgen`.


## Last updated
This vignette is associated with `AHgen` v1.0.0, and was last updated by Melissa Bedinger on `r Sys.Date()`.