---
title: "3 - Weighting edges"
author: 
 - Annie Visser-Quinn & Melissa Bedinger
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: paper
    highlight: haddock
vignette: >
  %\VignetteIndexEntry{3 - Weighting edges}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)

rm(list = ls()); cat("\014")

library(knitr)
library(tidyverse)
library(igraph)
library(AHgen)
```

<style>
#TOC {
  background: url("images/hex.png");
  background-size: 240px;
  background-position-x: center;
  background-position-y: top;
  padding-top: 280px !important;
  background-repeat: no-repeat;
}
</style>

```{r include = FALSE}
dh <- read_adjMat("../inputs/mobilityAH_adjMat_20230629.xlsx") 
dv <- dh %>% select(level, levelName_full, levelName, Node)
de <- dh %>% adjMat_to_edgelist(vInfo = dv)
di <- dh %>% adjMat_to_igraph(vInfo = dv)
```
---

<span style="color: red;">@Melissa Why you weight. What it means in the abstraction hierarchy context.</span>

In this vignette we'll explore **three methods of weighting edges** in R: 

- The first two are applied to the **edge list** and **igraph** directly using syntax from the `tidyverse` and `igraph` packages respectively
- The third way utilises the **AHgen functions** `read_weightScenario()` and `weightEdges()`


## 1. Weighting the edge list using `mutate()`

The `mutate()` function adds a new column of a length equal to the number of rows.

You can check the length of the edge list using `nrow()`. Here's an example for the mobility abstraction hierarchy:

```{r}
de %>% nrow()
```
This means that any new column must be of length 20, i.e. a vector of length 20. Now click on the examples below.

<details><summary><span style = "color: #2196F3;">**Show code - Example 1**</span></summary>

First, let's use random sampling to create a vector of length 20. 

```{r include = FALSE}
set.seed(1)
```
```{r}
vec <- sample(x = seq(0,1,by=0.01), size = 20)

vec %>% length

vec
```
We can now add this vector to the edge list using `mutate()` to replace the current *weight* column.

```{r}
de_w <- de %>% mutate(weight = vec)

de_w
```

</details>

<details><summary><span style = "color: #2196F3;">**Show code - Example 2**</span></summary>

Example 1 used randomly generated weights. In practice we would assign weights with some meaning. 

For example, an ageing vehicle. To reflect this, we could degrade the edges connecting to the vertex *Vehicle components*.

```{r}
de %>%
  mutate(weight = ifelse(from == "Vehicle components" | to == "Vehicle components", 0.75, weight)) %>%
  arrange(desc(layer)) 
```

This example would require prior coding knowledge to apply. The complexity would increase further if we wanted to change the weightings for multiple edges.

</details>




## 2. Weighting the igraph (using igraph syntax)
In vignette 1 we explored how to extract data and attributes from an igraph. Here's a quick recap.

```{r}
# Extract edges
E(di)

# Extract edge attribute weight
E(di)$weight
```

`E(di)$weight` is effectively an object. Therefore, to change the weighting, we can simply override the weight values using `<- vectorOfValues`. 

This vector of values must be the same length as the number of edges. From section 1 we know this is 20, but we can also check it using the igraph function that measures graph size, `gsize()`.

The examples below introduce two ways of weighting the igraph.

<details><summary><span style = "color: #2196F3;">**Show code - Example 3**</span></summary>

In this example, let's override the weights using `vec`, the vector created in Example 1.

```{r}
E(di)$weight <- vec
E(di)$weight
```

It's as simple as that!

</details>

```{r include = FALSE}
di <- dh %>% adjMat_to_igraph(vInfo = dv)
```

<details><summary><span style = "color: #2196F3;">**Show code - Example 4**</span></summary>

What if we want to change the weights of specific edges, like those connected to the *Vehicle components* vertex? We need to find out the index, that's the position, of the edge. The mobility abstraction hierarchy is small so this could easily be done by eye.

```{r}
E(di)[18:20]
```

All we'd have to do is overwrite this 'object' with the new weights. 

Let's imagine we don't know the index of the vertex, how can we find out? We begin by finding out the edge names.

```{r} 
edges <- E(di) [ from("Vehicle components") ]
edges
```
Now we need to find out the indices of these edges, we can do this using `which()` and `%in%`.

```{r}
index <- which(E(di) %in% edges)
index
```

Now we can overwrite the weightings.

```{r}
E(di)$weight[index] <- 0.75
E(di)$weight
```

As in example 2, this really require some prior coding knowledge to apply. It's also more difficult to visualise as the weights and edges can't be visualised side-by-side. The complexity would increase further if we wanted to change the weightings for multiple edges.
</details>

```{r include = FALSE}
di <- dh %>% adjMat_to_igraph(vInfo = dv)
```


## 3. Weighting in AHgen

```{r, include = FALSE}
genExample <- function(edgelist, layerName = NULL) { 
  
  set.seed(1)
  
  if(!is.null(layerName)) { edgelist <- edgelist %>% filter(layer == layerName) }
  
  output <- 
    edgelist %>%
    rowwise() %>% 
    mutate(weightNew = sample(seq(0.01,1,0.01),1:n())) %>%
    select(-weight)
  
  return(output)
  
}
  
de_w1 <- genExample(de, layerName = "l1FP_l2VPM")
de_w2 <- genExample(de, layerName = "l2VPM_l3GF")
```

Through examples 1-4 we've seen how to change weightings directly. These methods were:

- Cumbersome, requiring a degree of prior coding experience
- Impractical for cases where you want to weight edges connected to more than one vertex, i.e. the real world!

AHgen has two functions to simplify edge weighting which we'll explore now.


### 3.1 weightEdges()

The `weightEdges()` function has two input arguments: 

- *x* The edge list to be weighted
- *y* A weighted edge list with new weights for one or more edges in x

Below is an example of a weighted edge list which captures the edges in the *l1FP_l2VPM* layer only. With the `weightEdges()` function we can apply these weightings to our edge list.
```{r}
de_w1
```

When applying the function:

- Both edge lists must have columns named *layer*, *from* and *to*
- The weight column in the weighted edge list must be called *weightNew*
- The vertex names must be identical in both edge lists

**Most importantly, the weighted edge list can be generated in R or read in externally from a .csv or .xlsx file.**

Let's explore through some examples.

<details><summary><span style = "color: #2196F3;">**Example 5**</span></summary>
 
Let's start with the edge list shown above.

```{r}
weightEdges(de, de_w1)
```

It's that easy.

Let's look at another weighted edge list.

```{r}
de_w2

weightEdges(de, de_w2)
```
</details>

<details><summary><span style = "color: #2196F3;">**Example 6**</span></summary>

Weightings may be determined externally and imported. Let's revisit our case where a vehicle has been degraded due to age. The *mobilityAH_edges_exampleWeights_20230629s.xlsx* spreadsheet has been provided in the *inputs* folder. 

```{r}
de_w <- read_xlsx("../inputs/mobilityAH_edges_exampleWeights_20230629.xlsx")
de_w
```

Now that we've got our weighted edge list we can apply the weighting using `weightEdges()` as before.

```{r}
weightEdges(de, de_w)
```

Some handy tips:

- Export the edge list as a .csv to simplify editing, `de %>% write_csv("filepath.csv")`
- Import a csv using read_csv("filepath.csv", col_types = cols())

</details>


<span style="color: red;">Paused here</span>
## 3.2. Weighting using the indicators database

The Urban Systems Abstraction Hierarchy indicators database provides location specific information for a range of UK cities. These indicators are converted into weightings using the function `read_indicatorsDatabase()`. Input arguments are the database filepath and the city/location name. Located in the *inputs* folder (*indicatorsDatabase_2020-05-22.xlsx*), the indicators database is available in spreadsheet format for easy navigation. The spreadsheet may be adjusted or expanded. <span style="color: purple;">Provide more information???</span>

```{r}
# read_indicatorsDatabase("../inputs/indicatorsDatabase_2020-05-22.xlsx", "Bristol")
# read_indicatorsDatabase("../inputs/indicatorsDatabase_2020-05-22.xlsx", "Edinburgh")
# read_indicatorsDatabase("../inputs/indicatorsDatabase_2020-05-22.xlsx", "Glasgow")
```

These weightings can be applied using `weightEdges()`.

## Last updated
This vignette is associated with AHgen v1.0, and was last updated by Melissa Bedinger on 2023-07-03.