---
title: "AHgen - Vignette 3, Analysis"
author: "Dr Annie Visser-Quinn, a.visser-quinn@hw.ac.uk"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{AHgen - Vignette 3, Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(
  root.dir = normalizePath('../AHgen')
)

rm(list = ls()); cat("\014")

library(tidyverse)
library(igraph)
```

```{r, include = FALSE}
getwd()

load("envP2.RData")

dw_weighted <- weightLinks(x = di, y = dw_igraph, type = "igraph")

source("functions/functions.R")
```

This vignette follows on directly from *Vignette 2, Link weighting*. You should have the following objects loaded in your environment:

```{r}
ls()
```


## 1. Introduction
<span style="color: red;">Explain what type of analysis is done for abstraction hierarchy. Explain centralities (as a minimum).</span>


## 2. Metrics

### 2.1 Betweenness centrality
AHgen has two functions for calculating the unweighted and weighted betweenness centralities, `calcUWVBC()` and `calcWVBC()` respectively. There are two input arguments, an igraph and the node information (saved as *dn* in part 1).

Calculating the weighted vertex betweenness centrality requires the packages tnet and survival to be installed; `install.packages(c("tnet", "survival"))`.

```{r}
calcUWVBC(dw_weighted, dn)

calcWVBC(dw_weighted, dn)
```

### 2.2 Calculate multiple metrics

Multiple metrics can be calculated at the same time using the function `calcMetrics()`. The additional input argument is a vector of the names of the metrics to be calculated. Only functions which are described in `calcMetrics()` can be called; currently AHgen includes the unweighted and weighted vertex betweenness centralities only.

```{r}
calcMetrics(dw_weighted, dn, c("UWVBC", "WVBC"))
```

### 2.3 Adding additional metrics
Functions to calculate additional metrics can be easily added to AHgen by following the same structure as the function `calcUWVBC()`.

```{r, echo = FALSE}
calcUWVBC
```
The outputs from the igraph functions `betweenness()` and `eigen_centrality()` are the same format. Therefore we can create a function to calculate eigen centrality by changing three things in the code. If the new function's name is in the form *calcSuffix*, then `calcMetrics()` will be able to call it.

```{r}
calcEigen <- function(igraph, nodeInfo) { 
  
  require(igraph)
  
  igraph::eigen_centrality(igraph)$vector %>%
    function_igraphResultFormatting(name = "Closeness") %>%
    function_mergeNodeInfo(nodeInfo)
  
}

calcEigen(dw_weighted, dn)

calcMetrics(dw_weighted, dn, c("WVBC", "Eigen", "UWVBC"))

```



```{r}
function(igraph, nodeInfo, metrics) {
  
  output <- 
    
    lapply(metrics, function(x) {
    
    calc <- get(paste0("calc", x))
    
    calc(igraph = igraph, nodeInfo = nodeInfo)
    
  })
  
  output <- 
    output %>% 
    reduce(full_join, by = c("layer", "layerName", "nodeName")) %>%
    arrange(layer, layerName, nodeName)
  
  return(output)
  
}
```


### 2.4 Calculate change signal
To calculate the percent change from one igraph to another, for any given metric, we use the function `calcChange()`. There are three input arguments:

- *before* The metric outputs from the initial igraph
- *after* The metric outputs from the modified or weighted igraph
- *metric* The column name of the metric for which change is to be calculated

```{r}
calcChange(before = calcWVBC(di,dn), 
           after = calcWVBC(dw_weighted, dn), 
           metric = "WVBC")
```
 
The final vignette introduces potential ways of visualising the results and the network connections.