---
title: "AHgen - Vignette 4, Visualisation"
author: "Dr Annie Visser-Quinn, a.visser-quinn@hw.ac.uk"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{AHgen - Vignette 3, Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE,
                      fig.width = 10,
                      fig.height = 7)

knitr::opts_knit$set(
  root.dir = normalizePath('../AHgen')
)

rm(list = ls()); cat("\014")

library(tidyverse)
library(igraph)
```

```{r, include = FALSE}
getwd()

load("envP3.RData")

# source("functions/functions.R")
```

This vignette follows on directly from *Vignette 3, Analysis*. You should have the following objects loaded in your environment:

```{r}
ls()
```

## 1. Box plots
One way to visualise the change signal is boxplots. This example builds on the code used previously, as well as some tips from Information is Beautiful.

First, we need to save the data that we want to work with as an object.

```{r}
dc <- calcChange(before = calcWVBC(di,dn), 
                 after = calcWVBC(dw_weighted, dn), 
                 metric = "WVBC")
dc
```

Next we need to prepare the data into a format for plotting. We need to have the numbers for the boxplot as well as the axis text. `ggplot` has an inbuilt function that will generate the boxplot numbers, but this creates a *Tukey boxplot*. This type of boxplot is often used but it is more heavily focussed on outliers. In this example we produce a boxplot covering the full distribution of values by determining the minimum and maximum values, the interquartile range, and the median.

```{r} 
# Prepare the layers text
dc <- dc %>% mutate(names = paste0(layer, "-", layerName) %>% factor %>% fct_rev)

dc$names %>% unique
```

```{r}
# Prepare the numbers for the box plot
dcSummary <- 
  dc %>% 
  group_by(names) %>%
  summarise(min = min(pctChange),
            max = max(pctChange),
            lower = quantile(pctChange, 0.25),
            upper = quantile(pctChange, 0.75),
            median = median(pctChange),
            .groups = "drop")

dcSummary
```

```{r}
# Prepare the basic plot
dcSummary %>%
  ggplot(aes(x = names, ymin = min, lower = lower, middle = median, upper = upper, ymax = max)) +
  geom_errorbar() +
  geom_boxplot(stat = "identity")
```

```{r}
# Add some specifications for visuals
p <- 
  dcSummary %>%
  
  ggplot(aes(x = names, ymin = min, lower = lower, middle = median, upper = upper, ymax = max)) +
  geom_errorbar(width = 0.25, colour = "#212F3C") +
  geom_boxplot(stat = "identity", width = 0.5, colour = "#212F3C") +
  
  coord_flip() +
  labs(x = "Level of Abstraction Hierarchy", y = "Percent change") +
  
  theme_minimal(base_size = 11)

p
```

```{r}
# Once we're happy with the underlying plot we can make a few tweaks using theme. 
p <- 
  p + 
  theme(panel.grid.minor = element_blank(),
        panel.grid = element_line(colour = "#EBEDEF"),
        text = element_text(colour = "#85929E"),
        axis.text = element_text(colour = "#85929E"))

p
```

```{r}
# Overplotting points - why it's not a great idea

# This example is misleading as all the points will bunch up
# It's not clear how many points are located where
p + 
  geom_point(data = dc, aes(x = names, y = pctChange), inherit.aes = FALSE,
             colour = "#212F3C", size = 3, shape = 1) 

# This version is impossible to read, the points mask the boxplot
p +
  geom_jitter(data = dc, aes(x = names, y = pctChange), inherit.aes = FALSE,
             colour = "#212F3C", size = 3, shape = 1,
             height = 0, width = 0.25)
```

## 2. Ridge plot using ggridges
This example uses the R package `ggridges`. A ridge plot is a series of stacked density (PDF) plots. 

```{r}
library(ggridges)

dc %>% ggplot(aes(x = pctChange, y = names, fill = names)) + ggridges::geom_density_ridges()
```

```{r}
# With formatting
library(wesanderson)

dc %>% 
  ggplot(aes(x = pctChange, y = names, fill = names)) + 
  
  ggridges::geom_density_ridges(size = 0.5, colour = "#212F3C") +
  
  labs(y = "Level of Abstraction Hierarchy", x = "Percent change") +
  
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        panel.grid = element_line(colour = "#EBEDEF"),
        text = element_text(colour = "#85929E"),
        axis.text = element_text(colour = "#85929E"),
        legend.position = "none") +
  
  scale_fill_manual(values = wes_palette("Darjeeling1", n = 5)[c(1,4,3,5,2)])
```

## 3. Adjacency matrix plots
```{r}
# Split the template edge list by layer

de_list <- de %>% mutate(from = fct_inorder(from), to = fct_inorder(to)) %>% split(., .$layers)
de_list

p1 <- 
  de_list$l1FP_l2VPM %>% 
  rename(y = from, x = to) %>% 
  
ggplot(aes(x = str_wrap(x, width = 20), 
           y = str_wrap(y, width = 20))) + 
  
geom_tile(fill = "#85929E", colour = "#EBEDEF", size = 1) +
  
scale_x_discrete(expand = c(0,0)) +
scale_y_discrete(expand = c(0,0)) +
  
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        panel.grid = element_blank(),
        text = element_text(colour = "#85929E"),
        axis.text.y = element_text(colour = "#85929E"),
        axis.text.x = element_text(colour = "#85929E", angle = 90, hjust = 1, vjust = 0.5),
        axis.ticks = element_blank()) +
  
  labs(x = "Values and Priority Measures", y = "Functional purposes")

p1
```

```{r, fig.height = 10}
# We can reuse this plot using `%+%` and changing the labels

p2 <- p1 %+% (de_list$l2VPM_l3GF %>% rename(x = from, y = to)) + labs(y = "Generalised Functions")
p2
```
```{r, fig.height = 10}
# Repeat for the third level - becomes hard to read at this point
p3 <- p2 %+% (de_list$l3GF_l4ORP %>% rename(x = to, y = from)) + labs(x = "Object-related processes")
p3
```
An alternative way to do the adjacency matrix plot is to remove more visual features and make the plot interactive using the package `plotly`.

```{r}
# Use the same level of the AH, but add some weightings to try to make the visualisation more meaningful.
dt3 <- 
  de_list$l3GF_l4ORP %>% 
  rename(x = to, y = from) %>%
  mutate(weight = sample(seq(0,1,0.001), n()),
         Link = paste0("from ", y, " to ", x))
  
dt3
```
```{r, warning = FALSE, message = FALSE, fig.height = 12, fig.width = 12}
p4 <- 
  dt3 %>% 
  
  ggplot(aes(x = str_wrap(x, width = 20), 
           y = str_wrap(y, width = 20),
           label = Link)) + 
  
  geom_tile(aes(fill = weight), colour = "#EBEDEF", size = 1) +
  
  scale_x_discrete(expand = c(0,0)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_fill_viridis_c(option = "A", direction = -1) +
  
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        panel.grid = element_blank(),
        text = element_text(colour = "#85929E"),
        axis.text = element_blank(), 
        axis.ticks = element_blank()) +
  
  labs(x = "Generalised Functions", y = "Object-related processes")
p4

library(plotly)
ggplotly(p4, tooltip = "label")


```

Not included, but can be discussed, time allowing:

- Other types of matrix plot
- Fabric
- Focus network plots

<p>

Omitted from this version of AHgen:

- Using geospatial data
- Cutting out (flood) hazards
- Sensitvitiy