---
title: "AHgen - Vignette 2, Link weighting"
author: "Dr Annie Visser-Quinn, a.visser-quinn@hw.ac.uk"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{AHgen - Vignette 2, Link weighting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(
  root.dir = normalizePath('../AHgen')
)

rm(list = ls()); cat("\014")

library(tidyverse)
library(igraph)
```

```{r, include = FALSE}
getwd()

load("envP1.RData")

di <- de %>% edgelist_to_igraph(outputList = FALSE)
```

This vignette follows on directly from *Part 1 - Getting started*. You should have the following objects loaded in your environment:

```{r}
ls()
```


## 1. Introduction
<span style="color: red;">Is all the written material going to talk in terms of edges and links?</span>

<span style="color: red;">Add a line to explain why you weight - what it means in the abstraction hierarchy context. Normal terminology for networks is too generic.</span>

## 2. Weighting links

In igraph, links can be weighted in a number of ways. The simplest is illustrated in the code chunk below.

```{r}
E(di)$weight %>% head(n = 25) # Print first 25 weights

E(di)$weight[1:10] # Select elemnts 1 to 10

E(di)$weight[1:10] <- 0.5 # Change weightings of elements 1 to 10 to 0.5

E(di)$weight %>% head(n = 25)

E(di)$weight <- 1 # Change weightings back

E(di)$weight %>% head(n = 25)
```
The AHgen function `weightLinks()` is a user-friendly function for applying link weightings in the edge list or igraph format. There are three input arguments:

- *x* The edge list or igraph to be weighted
- *y* The edge list or igraph with the weights
- *type* The format of the objects, either *"edgelist"* or *"igraph"*

The link names in the weighted object must match the object to be weighted. The weightings edge list or igraph may be of any length - see example below.

To illustrate we will read in the csv file *weightExample* from the *inputs* folder.

```{r}
dw_edgelist <- read_csv("inputs/weightingExample.csv", col_types = cols())
dw_edgelist

dw_igraph <- dw_edgelist %>% rename(weight = weightNew) %>% edgelist_to_igraph
dw_igraph
```

### 2.1 Edge list
```{r}
weightLinks(x = de, y = dw_edgelist, type = "edgelist")
```

### 2.2 igraph
```{r}
dw_weighted <- weightLinks(x = di, y = dw_igraph, type = "igraph")
dw_weighted

E(dw_weighted)$weight %>% head(n = 25)
```

## 3. Indicators database

<span style="color: red;">May need to provide a bit more background information.</span> The indicators database provides location specific information, for a range of UK cities. These indicators are converted into weightings using the function `read_indicatorsDatabase()`. Input arguments are the database directory & name as well as the name of the city / location.

Located in the *inputs* folder (*indicatorsDatabase_2020-05-22.xlsx*), the indicators database is available in spreadsheet format for easy navigation. The spreadsheet may be adjusted, expanded or copied for use in AHgen.

<span style="color: red;">Provide more information on how the weighting values are calculated</span>

<span style="color: green;">Get Melissa to confirm that the method of calculation is correct.</span>. 

```{r}
read_indicatorsDatabase("inputs/indicatorsDatabase_2020-05-22.xlsx", "Bristol")
read_indicatorsDatabase("inputs/indicatorsDatabase_2020-05-22.xlsx", "Edinburgh")
read_indicatorsDatabase("inputs/indicatorsDatabase_2020-05-22.xlsx", "Glasgow")
```

<p>

These weightings can then be applied to the generic template using `weightLinks()`.
```{r}
id <- read_indicatorsDatabase("inputs/indicatorsDatabase_2020-05-22.xlsx", "Edinburgh")

weightLinks(de, id, "edgelist")
```

<p>

These indicator weightings can also be applied to the level of the abstraction hierarchy directly.

```{r}
de_list$l2VPM_l3GF

de_list$l2VPM_l3GF <- weightLinks(de_list$l2VPM_l3GF, id, "edgelist")

de_list$l2VPM_l3GF
```

The walkthrough continues with **Part 3 - Analysis**.